---
title: "Specialty Distribution and Surgeon-Level Concentration of Sling Surgery for Stress Urinary Incontinence in Fee-for-Service Medicare, 2013–2023"
author: "Tyler Muffly, MD"
output:
  word_document:
    reference_docx: null
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressWarnings(suppressMessages({
  library(dplyr); library(knitr); library(readr)
}))
# Supplementary tables and figures are produced by pipeline steps 06 and 07
# from the same cache, so they stay consistent with the inline numbers. Both
# helpers degrade gracefully if an artifact is missing.
# Paths are relative to knit_root_dir (the project root), which is where chunk
# *code* runs. Guards live inside chunk code, not in the `eval` option, because
# knitr evaluates chunk options relative to the Rmd's own folder (output/).
fig_dir    <- "output/figures"
tables_dir <- "output/tables"
# Emit a raw-markdown image with an ABSOLUTE path. Pandoc resolves absolute
# paths regardless of its working directory, so the PNG always embeds into the
# docx. (include_graphics() records a root-relative path that pandoc, run from
# the Rmd's folder, cannot find, so it silently drops the image.)
show_fig <- function(name, width = "6.5in") {
  p <- normalizePath(file.path(fig_dir, name), mustWork = FALSE)
  if (file.exists(p)) cat(sprintf("\n![](%s){width=%s}\n\n", p, width))
  else cat(sprintf("\n_Figure not found (%s); run pipeline step 07._\n\n", name))
}
read_supp <- function(name) {
  p <- file.path(tables_dir, paste0(name, ".csv"))
  if (!file.exists(p)) return(NULL)
  suppressWarnings(readr::read_csv(p, show_col_types = FALSE))
}
fmt_pcol <- function(x) {
  x <- suppressWarnings(as.numeric(x))
  ifelse(is.na(x), "n/a", ifelse(x < 0.001, "<0.001", sprintf("%.3f", x)))
}
# Every number below is computed from the pipeline cache, not hardcoded, so the
# prose can never drift from the analysis. The path defaults to the pipeline
# cache and can be overridden with the PUF_CLASSIFIED environment variable.
source("R/compute_manuscript_values.R")
cache_dir <- tryCatch(config::get("cache_dir"), error = function(e) "data/cache")
puf_path  <- Sys.getenv("PUF_CLASSIFIED", file.path(cache_dir, "puf_classified.rds"))
mv  <- compute_manuscript_values(puf_path)
v   <- mv$v
tab <- mv$tab
kt  <- function(df, align = NULL) knitr::kable(df, align = align, format = "pipe")
# composite inline helpers
iqr  <- function(m, lo, hi) sprintf("%g (%g–%g)", m, lo, hi)
ci   <- function(s, lo, hi, p) sprintf("%+.2f percentage points/year; 95%% CI %.2f to %.2f; p %s",
                                       s, lo, hi, ifelse(p == "<0.001", "< 0.001", paste0("= ", p)))
```

**Author:** Tyler Muffly, MD^1^

^1^ *[Affiliation to be completed: division, department, institution, city, state]*

**Corresponding author:** Tyler Muffly, MD. *[mailing address, telephone, and email to be completed]*

**Funding:** *[to be completed; state "None" if not applicable]*

**Conflicts of interest:** *[to be completed; state "The author reports no conflicts of interest" if applicable]*

**Acknowledgments:** *[to be completed]*

**Data and code availability:** The complete, reproducible analytic pipeline and this manuscript source are available at https://github.com/mufflyt/sling-volume-patterns. The CMS Medicare Physician and Other Practitioners Public Use File and the CMS Program Statistics Original Medicare Enrollment tables are publicly available from data.cms.gov.

**Word count (abstract):** approximately 250. **Manuscript type:** Original research.

---

## Abstract

**Objectives:** To describe the specialty distribution, annual clinician volume, and surgeon-level concentration of sling surgery for stress urinary incontinence (CPT 57288) in fee-for-service Medicare from 2013 through 2023.

**Methods:** We analyzed all 11 annual CMS Medicare Physician and Other Practitioners Public Use Files. Individual clinicians were classified by CMS provider type and ABOG and ABU roster linkage. We modeled utilization with female Part B fee-for-service enrollment as an offset, annual volume with a Poisson generalized estimating equation clustered by NPI, and URPS share with a quasibinomial model. Annual Gini coefficients measured concentration.

**Results:** Among `r format(v$analytic_physicians, big.mark=",")` clinicians, `r format(v$analytic_procs, big.mark=",")` services were observable. OB/GYN-pathway URPS accounted for `r v$urps_pct`% of services and urology-pathway URPS for `r v$urpsuro_pct`%, for a combined URPS share of `r v$urps_combined_pct`%. Reported services declined `r format(abs(v$count_pct_change), nsmall=1)`%; after enrollment adjustment, the estimated utilization rate declined `r sub("^-","",v$rate_offset_pctyr)`% per year (rate ratio `r v$rate_offset_rr`, 95% CI `r v$rate_offset_ci`; p = `r v$rate_offset_p`). OB/GYN-pathway URPS had the highest adjusted annual volume. Within-year concentration remained stable. The fitted OB/GYN-pathway URPS share increased from `r v$share_binom_2013`% to `r v$share_binom_2023`% (+`r v$share_binom_pp` percentage points/year; p `r ifelse(v$share_binom_p=="<0.001","< 0.001",paste("=",v$share_binom_p))`), whereas the urology-pathway share changed comparatively little.

**Conclusions:** In observable fee-for-service Medicare, URPS physicians performed most sling services, and their share increased without greater surgeon concentration. These findings describe claims-visible service patterns, not the complete national surgical workforce; all-payer outcome and access studies are needed before drawing credentialing implications.

**Keywords:** stress urinary incontinence; sling surgery; urogynecology; Medicare; surgical workforce; procedural concentration.

---

## Simply Stated

Sling surgery treats stress urinary incontinence, which causes urine leakage with coughing, laughing, or exercise. We studied traditional Medicare billing records from 2013 through 2023 to learn which doctors performed these operations and how the work was shared. Urogynecologists performed most of the reported surgeries, and their share increased over time. The total number of reported surgeries fell. Some of that decline occurred because fewer people were enrolled in traditional Medicare, although a smaller true decline may remain. The work stayed spread across many surgeons rather than becoming centered in a small group. These data cannot show which specialty has the best outcomes or who should be allowed to perform the operation.

## Why This Matters

National data on who performs sling surgery are limited. This study improves classification by identifying urogynecologists through both the obstetrics-gynecology and urology certification pathways, avoiding the misclassification of urology-trained urogynecologists as general urologists. It also pairs claims with fee-for-service enrollment, showing that raw service counts overstate the decline in utilization. The main finding is a growing URPS share without increasing within-year surgeon concentration. These results support workforce surveillance, but claims data alone cannot establish quality, access, or credentialing standards.

---

## Introduction

Sling surgery is a common operation for stress urinary incontinence (SUI) and remains a standard surgical treatment when conservative care fails.^1,2^ These operations are reported under CPT 57288, whose descriptor is "sling operation for stress incontinence (eg, fascia or synthetic)." The code therefore captures autologous fascial and synthetic sling operations together and does not identify the sling material or distinguish retropubic, transobturator, single-incision, or pubovaginal technique. Despite the procedure's widespread use, little is known about which specialties perform these operations in the Medicare population or how volume is distributed among surgeons.

These questions matter for surgical quality, training, and workforce planning. Across surgery, higher surgeon and hospital volume is generally associated with better outcomes.^4,5^ Evidence for procedure-specific volume thresholds in sling surgery is more limited. The specialty mix has also changed during the past decade.

Three developments make this analysis timely. First, Female Pelvic Medicine and Reconstructive Surgery, now named Urogynecology and Reconstructive Pelvic Surgery (URPS), has grown since receiving American Board of Medical Specialties recognition in 2013.^6^ Certification is available through both obstetrics and gynecology (ABOG) and urology (ABU) pathways. Second, US Food and Drug Administration actions on urogynecologic mesh changed public perception of mesh and may have affected sling use.^7^ The 2019 FDA order concerned transvaginal mesh for pelvic organ prolapse and did not remove synthetic midurethral slings for stress incontinence from the market; the two devices are frequently conflated in public discussion. Third, procedural concentration, the extent to which a relatively small group of clinicians performs a large share of procedures, has become relevant to quality improvement and resource allocation.^8^

We therefore characterized specialty distribution, annual surgeon volume, and surgeon-level concentration among physicians performing sling surgery for stress urinary incontinence in fee-for-service Medicare from 2013 through 2023. We also evaluated changes in each specialty's share of reported services over time.

## Materials and Methods

### Study Design and Data Source

We conducted a repeated cross-sectional analysis of the Centers for Medicare & Medicaid Services (CMS) Medicare Physician and Other Practitioners Public Use File (PUF) from 2013 through 2023. The PUF contains all fee-for-service Medicare Part B claims aggregated to the provider-service level. Each row represents a unique combination of National Provider Identifier (NPI), Healthcare Common Procedure Coding System code, place of service, and calendar year. Because the dataset is publicly available and contains no protected health information, institutional review board approval was not required.

All 11 annual PUF releases (2013 through 2023) were verified against the CMS file specifications for completeness before analysis.

### Specialty Classification

Providers were assigned to four mutually exclusive groups: URPS, MIGS, General OB/GYN, or non-URPS urology. We first used the annual CMS provider type (Rndrng_Prvdr_Type), which may vary by year, to identify OB/GYN, urology, or another specialty. Among OB/GYN physicians, the ABOG subspecialty registry (a snapshot of ABOG-certified diplomates linked to NPIs primarily by exact NPI match and, when needed, by full name and practice state) identified "Female Pelvic Medicine and Reconstructive Surgery" as URPS and "MIG" as MIGS; all other OB/GYN physicians were classified as General OB/GYN. This yielded `r v$analytic_physicians`-cohort assignments of URPS, MIGS, and General OB/GYN through the OB/GYN pathway.

Billers of CPT 57288 who had neither an OB/GYN nor urology CMS provider type and were not in the ABOG registry (`r v$class_reclass_urology` NPIs) were predominantly facilities (ambulatory surgical centers) and non-physician clinicians (physician assistants, nurse practitioners), not urologists; none matched the ABU roster. We therefore did not assign them to urology. Organizational NPIs (NPPES entity type 2, flagged in the Public Use File by a rendering-provider entity code of "O": ambulatory surgical centers, hospitals, laboratories, and groups; `r v$class_excluded_facility` NPIs) were excluded because they identify a billing organization rather than an individual clinician and cannot be given a physician specialty. The remaining `r v$other_phys` individual non-physician or unclassifiable clinicians were placed in a distinct "Other/uncertain" group (`r v$other_pct`% of observable services). The full classification flow, with counts at each step, is shown in Figure 3. We report the specialty distribution under three handlings of these billers, as a sensitivity analysis (Supplementary Table S11): the primary "Other/uncertain" grouping, exclusion, and the legacy assignment to urology; the latter inflates the apparent non-URPS urology share roughly two-fold and is not used. A further small number of provider-year records from ABOG-registered NPIs with a miscellaneous CMS type remained "Other" and were excluded.

We then cross-referenced the American Board of Urology roster (`r nrow(readr::read_csv("data/abu_urology/abu_urps_npi_LATEST.csv", show_col_types=FALSE))` urology-pathway urogynecologist NPIs) to identify urology-pathway urogynecologists; `r v$class_abu_pathway` cohort physicians matched. These physicians were combined with ABOG-certified urogynecologists into a single all-pathway URPS group; all remaining urologists were classified as non-URPS urology. Without this step, fellowship-trained urology-pathway urogynecologists would appear only as "Urology" in the PUF. Because an all-pathway URPS group includes urology-trained subspecialists who are not gynecologists, we report the all-pathway URPS share separately from the OB/GYN-based share (URPS plus MIGS plus General OB/GYN, restricted to the OB/GYN training pathway), rather than a single "gynecologic" share.

Specialty assignment could affect the estimated market-share trend because 45% of URPS physicians in the cohort were certified after 2013. We therefore examined two plausible classification scenarios that make opposite assumptions about pre-certification practice years. Fixed-membership classification assigns each physician's eventual subspecialty to every study year; because it counts physicians as URPS before certification, it tends to produce a shallower estimated slope.

Time-varying, certification-gated classification counts a physician as URPS or MIGS only beginning in the ABOG subspecialty certification year (sub1startdate, 2013–2024, distinct from the initial OB/GYN board date). Earlier years are classified by that year's CMS provider type. Because certification generally follows the start of subspecialty practice (the 2013 examination certified physicians already practicing urogynecology, and later diplomates also practiced before certification), this scenario removes some genuine pre-certification practice from the early URPS count and tends to produce a steeper estimated slope. We present these as plausible scenarios rather than formal statistical bounds, because neither assumption is guaranteed to bracket the true trend.

Modal and ever-URPS/MIGS classifications were additional sensitivity analyses. Urology-pathway URPS classification remained fixed because no urology subspecialty certification date was available. For the whole-period distribution, concentration, and volume analyses (Tables 1–3), each physician was assigned to the eventual combined subspecialty.

### CMS Data Suppression

CMS suppresses provider-service data when a provider treats fewer than 11 Medicare beneficiaries for a given code in a calendar year. Every observable provider in this study therefore treated at least 11 Medicare beneficiaries for CPT 57288 that year. Providers below the threshold are not observed, so the true number of low-volume surgeons is unknown. Suppression removes the lowest-volume physician-years and can bias the Gini coefficient and HHI in different directions, so the full-market concentration cannot be identified from the Public Use File. We therefore report bootstrap confidence intervals for the Gini coefficient and a sensitivity analysis that adds hypothetical suppressed low-volume providers before recomputing concentration.

### Outcome Measures and Statistical Analysis

The primary surgeon-level outcome was the annual number of CPT 57288 services per provider (Tot_Srvcs, reported services). To describe population utilization over time, we also computed the annual rate of CPT 57288 services per 100,000 female Part B fee-for-service Medicare beneficiaries, using the CMS Program Statistics Original Medicare Enrollment counts (table MDCR ENROLL AB 11, Sex = Female, Part B) as the denominator; absolute service counts are reported as a secondary measure. Because the Public Use File records no beneficiary age for the numerator, the rate is crude and is not age-standardized. We measured surgeon-level concentration with complementary metrics. The Gini coefficient summarizes inequality across the full surgeon-volume distribution, and we bootstrapped its 95% confidence interval by resampling providers. The Herfindahl–Hirschman Index (HHI) sums squared surgeon shares on a 0–10,000 scale and is more sensitive to the largest-volume surgeons. Because the minimum attainable HHI depends on the number of providers, we do not compare raw HHI directly across groups of very different size (URPS `r v$urps_phys`, General OB/GYN `r v$gob_phys`, MIGS `r v$mig_phys` physicians); we also report the size-adjusted normalized HHI and the effective number of providers (1 divided by the sum of squared shares). Within-year concentration, computed separately for each calendar year, was the primary concentration measure; pooled multi-year values are secondary because they mix annual volume with how many years each physician remained observable.

The physician was the operative production unit in these calculations. Thus, HHI describes surgeon-level procedural concentration rather than hospital or health-system market competition and should not be interpreted using FTC/DOJ antitrust thresholds.

Because each physician contributed as many as eleven annual observations, physician-year records were correlated. We modeled the annual count of reported services, conditional on the physician-year being observable (above the CMS suppression threshold), with a Poisson generalized estimating equation clustered by NPI, using an exchangeable working correlation and robust standard errors. Fixed effects were specialty (URPS reference), calendar year centered at 2018, a specialty-by-year interaction, and a 2020 COVID indicator. Centering at mid-study makes the specialty main effects the adjusted rate ratios at 2018; because the interaction makes the year term specialty-specific, we report each specialty's annual slope as a marginal linear contrast (the year term plus its interaction) rather than a single overall time effect. We report adjusted rate ratios with 95% CIs. A negative-binomial mixed model with a random intercept for NPI was fit as a sensitivity analysis, and its estimates are reported alongside the GEE. Additional sensitivity analyses excluded calendar year 2020 and restricted the cohort to physicians observable in at least two years.

As a secondary analysis with one independent observation per physician, we compared each physician's median annual volume across specialties using the Kruskal-Wallis test and Bonferroni-adjusted pairwise Wilcoxon tests.

To estimate the URPS market-share trend, we modeled URPS services as a proportion of all annual services with a quasibinomial generalized linear model on calendar year (centered at 2018), which respects the compositional structure that separate ordinary least-squares regressions on annual percentages ignore. We express the trend as an odds ratio per year and as the marginal percentage-point change per year. Separate ordinary least-squares regressions of each specialty's annual percentage share (URPS, urology, General OB/GYN) are reported as a descriptive sensitivity. MIGS was described without formal emphasis because only `r v$mig_phys` physicians were identified.

To describe observable participation transitions, we used a two-year washout. A surgeon was counted as newly observable in a given year when present but absent in both prior observable years; as continuing when present and not newly observable; and as no longer observable when absent in both subsequent observable years. We tabulated the share of annual volume performed by newly observable surgeons, their median volume, and their distribution by specialty. Because CMS suppression removes providers below 11 beneficiaries, these transitions describe when a surgeon crosses the reporting threshold, not genuine entry into or exit from practice; a surgeon may become unobservable and later reappear. Counts of newly observable surgeons by specialty are episode counts (a physician observed as newly observable in more than one year, after an intervening gap, contributes more than once), not counts of unique physicians.

In a secondary geographic analysis, we tabulated observable surgeons and URPS share by practice state and identified states with no observable URPS surgeon performing at least 11 Medicare slings in any study year. We did not calculate state-level population-based rates because state-by-sex fee-for-service enrollment denominators were outside the scope of this analysis; the national utilization rate above uses the national female Part B fee-for-service denominator.

Analyses were performed in R 4.4, with package versions locked through renv. The complete analytic pipeline is available at https://github.com/mufflyt/sling-volume-patterns.

## Results

### Cohort and Annual Volume

Across the full 11-year window, `r format(v$full_physicians, big.mark=",")` physicians billed CPT 57288, accounting for `r format(v$full_services, big.mark=",")` observed services in `r format(v$full_pyears, big.mark=",")` physician-year observations. Participation was intermittent: `r v$persist_1` physicians (`r v$persist_1_pct`%) appeared in one year, `r v$persist_2_5` (`r v$persist_2_5_pct`%) in 2–5 years, `r v$persist_6_10` (`r v$persist_6_10_pct`%) in 6–10 years, and `r v$persist_11` (`r v$persist_11_pct`%) in all 11 years. Thus, only a small proportion maintained an observable Medicare sling practice throughout the study window.

The observed count of reported services fell from `r format(v$procs_2013, big.mark=",")` in 2013 to `r format(v$procs_2023, big.mark=",")` in 2023 (`r format(v$count_pct_change, nsmall=1)`%). Over the same period the female Part B fee-for-service population contracted from `r v$den_2013_m` million to `r v$den_2023_m` million beneficiaries (`r format(v$den_pct_change, nsmall=1)`%). In a Poisson model of the annual service count with log fee-for-service enrollment as an offset and a 2020 indicator, the observable utilization rate declined `r v$rate_offset_pctyr`% per year (rate ratio `r v$rate_offset_rr`, 95% CI `r v$rate_offset_ci`; about `r v$rate_offset_decade`% over the decade), but with only 11 annual observations and the pandemic interruption this trend was not statistically robust (dispersion-adjusted p = `r v$rate_offset_p`). The rate reached a low of `r v$rate_2020` per 100,000 in 2020, consistent with pandemic-related deferral of elective surgery, then partially recovered. The number of observable surgeons also fell, from `r v$surg_2013` to `r v$surg_2023`. A clinically meaningful decline in observable utilization thus remains possible, but the raw service-count drop overstates it because much of that drop reflects the shrinking fee-for-service denominator; because CMS suppresses low-volume physician-years, this is an observable-services rate, not the total national fee-for-service rate.

### Specialty Distribution and Trends

We separated URPS into its two certification pathways (Table 1). OB/GYN-pathway URPS (ABOG-certified) was the largest single group: `r v$urps_phys` physicians, `r format(v$urps_procs, big.mark=",")` services (`r v$urps_pct`%), median annual volume `r v$urps_med` (p25–p75, `r v$urps_p25`–`r v$urps_p75`). Urology-pathway URPS (ABU roster) added `r v$urpsuro_phys` physicians and `r format(v$urpsuro_procs, big.mark=",")` services (`r v$urpsuro_pct`%; median `r v$urpsuro_med`), so the two URPS pathways together performed `r v$urps_combined_pct`% of observable services. Non-URPS urology included `r v$uro_phys` physicians and `r format(v$uro_procs, big.mark=",")` services (`r v$uro_pct`%; median `r v$uro_med`), other non-URPS OB/GYN `r v$gob_phys` physicians and `r format(v$gob_procs, big.mark=",")` services (`r v$gob_pct`%; median `r v$gob_med`), the Other/uncertain group `r v$other_phys` clinicians (`r v$other_pct`%), and MIGS `r v$mig_phys` physicians (`r v$mig_pct`%; median `r v$mig_med`).

Because `r v$switchers` clinicians changed specialty groups across years, specialty-specific counts exceed the unique cohort total. The urology-pathway URPS group (`r v$abu_moved_phys` physicians, `r format(v$abu_moved_procs, big.mark=",")` services) would otherwise be classified as general urology, understating the subspecialty and overstating non-URPS urology.

Annual market shares moved in different directions (Table 1). URPS increased from `r v$urps_trend$s13`% in 2013 to `r v$urps_trend$s23`% in 2023 (`r ci(v$urps_trend$slope, v$urps_trend$lo, v$urps_trend$hi, v$urps_trend$p)`). Urology decreased from `r v$uro_trend$s13`% to `r v$uro_trend$s23`% (`r ci(v$uro_trend$slope, v$uro_trend$lo, v$uro_trend$hi, v$uro_trend$p)`), and General OB/GYN decreased from `r v$gob_trend$s13`% to `r v$gob_trend$s23`% (`r ci(v$gob_trend$slope, v$gob_trend$lo, v$gob_trend$hi, v$gob_trend$p)`). MIGS increased descriptively from `r v$mig_trend$s13`% to `r v$mig_trend$s23`% (+`r sprintf("%.2f", v$mig_trend$slope)` percentage points/year), but this estimate is based on only `r v$mig_phys` physicians. In the quasibinomial model of URPS services out of all annual services, which respects the compositional structure, the URPS share rose from a fitted `r v$share_binom_2013`% in 2013 to `r v$share_binom_2023`% in 2023 (odds ratio `r v$share_binom_or` per year; `r v$share_binom_pp` percentage points per year; p `r ifelse(v$share_binom_p=="<0.001","< 0.001",paste("=",v$share_binom_p))`), consistent with the descriptive ordinary least-squares estimate. Because total service counts declined, these changes represent redistribution among specialties rather than growth in services.

The URPS increase was significant under every classification scenario, which differed in the estimated slope (Table 4). Fixed membership gave the shallower estimate because it counts physicians as URPS before certification: the all-pathway URPS share rose from `r v$fixed_urps_s13`% to `r v$fixed_urps_s23`% (+`r sprintf("%.2f", v$fixed_urps_slope)` percentage points/year), and the OB/GYN-based share from `r v$gyn_fixed_s13`% to `r v$gyn_fixed_s23`% (+`r sprintf("%.2f", v$gyn_fixed_slope)` percentage points/year). Certification-gated classification gave the steeper estimate, from `r v$ct_urps_s13`% to `r v$ct_urps_s23`% (+`r sprintf("%.2f", v$ct_urps_slope)` percentage points/year), because it removes not-yet-certified physicians from the early URPS count; the two estimates converged near `r v$ct_urps_s23`% by 2023.

In the Poisson GEE clustered by NPI (reference = OB/GYN-pathway URPS; calendar year centered at 2018), adjusted annual volume at mid-study was lower than OB/GYN-pathway URPS for every other group, including urology-pathway URPS (RR `r v$gee_urpsuro`), non-URPS urology (RR `r v$gee_urology`), other non-URPS OB/GYN (RR `r v$gee_gob`), and MIGS (RR `r v$gee_migs`) (Table 3). Annual volume was approximately 11% lower in 2020 (RR `r v$gee_covid`; p `r ifelse(v$gee_covid_p=="<0.001","< 0.001",paste("=",v$gee_covid_p))`). The specialty-by-year interaction showed that per-physician volume was essentially flat over time for the well-populated groups (OB/GYN-pathway URPS RR `r v$slope_urps`, p = `r v$slope_urps_p`; urology-pathway URPS RR `r v$slope_urpsuro`, p = `r v$slope_urpsuro_p`; non-URPS urology RR `r v$slope_uro`, p = `r v$slope_uro_p`; other OB/GYN RR `r v$slope_gob`, p = `r v$slope_gob_p`), increasing only for the 10-physician MIGS group (RR `r v$slope_migs`; p = `r v$slope_migs_p`) (Table 3). A negative-binomial mixed model with a random intercept per NPI reproduced the direction but with attenuated magnitude (urology RR `r if (v$has_nb) v$nb_urology else "NA"`, General OB/GYN RR `r if (v$has_nb) v$nb_gob else "NA"`, 2020 RR `r if (v$has_nb) v$nb_covid else "NA"`); the attenuation is expected because the GEE estimates a population-averaged (marginal) rate ratio whereas the mixed model estimates a physician-conditional one. The one-observation-per-physician secondary analysis agreed (Kruskal-Wallis H = `r v$kw_H`, df = `r v$kw_df`, p `r ifelse(v$kw_p=="<0.001","< 0.001",paste("=",v$kw_p))`); URPS volume exceeded urology and other OB/GYN in pairwise comparisons (p < 0.001), with no significant URPS-MIGS difference. Because these models condition on the physician-year being observable above the CMS reporting threshold, the estimand is annual service volume among observable physician-years, not full-workforce practice volume; if non-URPS physicians more often fall below the threshold, the URPS volume advantage may be overstated.

### Surgeon Volume and Concentration

The primary concentration analysis was within-year. Within-year concentration was low and stable across the study period: the annual Gini ranged from `r v$annual_gini_lo` to `r v$annual_gini_hi`, with no temporal trend (p = `r v$annual_gini_p`), and the annual top 20% performed approximately `r v$annual_top20`% of cases. No specialty-specific annual Gini changed significantly (all p > 0.15). Thus, care did not become concentrated among a smaller group of high-volume surgeons as the observable surgeon pool contracted.

Pooled multi-year concentration was modest for every group: URPS Gini `r v$urps_gini_ci`, urology Gini `r v$uro_gini_ci`, and other non-URPS OB/GYN Gini `r v$gob_gini_ci` (Table 2). Because overlapping individual confidence intervals are not a test of between-group difference, we bootstrapped the pairwise differences: URPS and urology did not differ (Gini difference `r v$ginidiff_urps_uro$txt`), whereas URPS was modestly less concentrated than other OB/GYN (`r v$ginidiff_urps_gob$txt`, confidence interval excluding zero). The size-adjusted normalized HHI was near zero for every group (URPS `r v$urps_hhi_norm`, urology `r v$uro_hhi_norm`, other OB/GYN `r v$gob_hhi_norm` on a 0–1 scale). The effective number of providers, expressed as a fraction of the actual provider count, showed a gradient (URPS `r v$urps_effn` of `r v$urps_phys`, `r v$urps_effpct`%; urology `r v$uro_effpct`%; other OB/GYN `r v$gob_effpct`%), indicating that sling volume was distributed across many surgeons rather than dominated by a few, though somewhat less evenly in the smaller groups. Because the specialties differ several-fold in provider count, raw HHI is not directly comparable across them (Table 2). Adding hypothetical suppressed low-volume providers raised the Gini (Supplementary Table S8): under the 50% scenario the URPS Gini rose from `r sprintf("%.2f", v$urps_gini)` toward the mid-0.6 range, so the sensitivity supports only the narrower claim that observed services were not dominated by a handful of surgeons, not a precise estimate of the full-market distribution. URPS had the highest median annual volume and upper quartile (median `r v$urps_med`; p75 `r v$urps_p75`). MIGS was excluded from these comparisons because only `r v$mig_phys` physicians were identified (1–4 per year); its estimates are exploratory.

### Observable Participation Over Time

Using the two-year washout (Supplementary Table S9), `r v$ent_min`–`r v$ent_max` surgeons became newly observable each year and performed `r v$entpct_min`%–`r v$entpct_max`% of annual volume. Their median volume was low, at approximately `r v$ent_med` services, consistent with surgeons appearing just above the CMS suppression threshold. Newly observable counts fell to `r v$ent_2020` surgeons (`r v$entpct_2020`% of volume) in 2020, then rebounded to `r v$ent_2022` (`r v$entpct_2022`%) in 2022; the 2022 rebound may partly reflect re-observation of surgeons who fell below the threshold during the pandemic rather than genuinely new surgeons. Continuing surgeons (`r v$cont_min`–`r v$cont_max` per year) performed most annual services, while surgeons no longer observable ranged from `r v$exit_min` to `r v$exit_max` per year.

Across the study period, URPS accounted for the most newly observable episodes (`r v$ent_urps`), followed by urology (`r v$ent_uro`), General OB/GYN (`r v$ent_gob`), and MIGS (`r v$ent_migs`); these are episode counts, not unique physicians. The rising URPS share coincided with the largest inflow of newly observable surgeons together with a relatively stable continuing-URPS base (URPS surgeon count `r v$urps_surg_2013` to `r v$urps_surg_2023`; per-physician volume unchanged). Urology showed substantial turnover with net decline, and General OB/GYN declined in both observable surgeons and market share.

Because total service counts declined, the growth in URPS share represents a redistribution of a shrinking service pool toward URPS, not new procedure growth. These transitions describe crossings of the CMS reporting threshold, not definitive entry into or exit from practice.

Observable surgeons and the URPS share by practice state are reported as an exploratory secondary analysis in Supplementary Table S7. Because state-by-sex fee-for-service denominators were not available and CMS suppression removes low-volume providers, we do not draw access conclusions from these counts; population-based geographic analysis is reserved for a separate access-focused study.

## Discussion

This national Medicare analysis produced three main findings. First, OB/GYN-pathway URPS was the largest single group (`r v$urps_pct`%) and had the highest adjusted per-physician volume, and the two URPS pathways together performed the majority (`r v$urps_combined_pct`%) of observable services. Second, the OB/GYN-pathway URPS share increased over the decade, coinciding with more URPS surgeons becoming observable and fewer non-URPS surgeons remaining observable, while the urology-pathway URPS share was roughly stable. Third, surgeon-level concentration remained low and stable despite a shrinking observable surgeon pool. Together, these findings indicate a change in the composition of observable fee-for-service practice rather than intensification of individual practice or concentration of care; they are best read as a practice-pattern study, not a study of entry into or exit from the clinical workforce.

### Specialty Distribution in Context

The two URPS pathways together performed approximately three-fifths of observable fee-for-service Medicare slings (`r v$urps_combined_pct`%; OB/GYN pathway `r v$urps_pct`%, urology pathway `r v$urpsuro_pct`%), and non-URPS urologists a much smaller share (`r v$uro_pct`%). Because the all-pathway URPS group also includes urology-trained subspecialists, the URPS share is not equivalent to a gynecologist share. The apparent non-URPS urology share is sensitive to classification: a naive provider-type approach that assigns ambulatory surgical centers and non-physician billers to urology roughly doubles it (Supplementary Table S11), which likely explains why some earlier analyses reported a larger urology share. Direct comparison with prior studies is limited because they used different datasets, specialty definitions, and care settings. In ACS-NSQIP data from 2006 through 2013, James et al. reported that gynecologists performed 74.2% of sling procedures and urologists 25.8%;^9^ Cantrell et al. found urogynecologists performed 54% of stress-incontinence procedures at academic centers from 2009 through 2014;^10^ and in Medicare data from 2002 through 2007, Rogo-Gupta et al. found urologists performed most sling operations.^11^ Rather than asserting a single historical reversal across these heterogeneous sources, we conclude more cautiously that OB/GYN-pathway URPS was the largest single contemporary fee-for-service Medicare group performing this procedure.

Including both certification pathways was essential to this result. Cross-referencing the ABU roster reassigned `r v$abu_moved_phys` urology-typed physicians, representing `r sprintf("%.1f", 100*v$abu_moved_procs/v$analytic_procs)`% of all sling procedures, to URPS. The concentration estimate was robust: the pooled URPS Gini changed only from `r sprintf("%.2f", v$urps_gini)` to 0.52. A pathway-neutral definition also reflects differences in operative case volume between urology- and gynecology-based URPS fellowships^14^ and variable FPMRS exposure during urology residency.^15^ Classifying urology-pathway urogynecologists as general urologists would therefore understate the true subspecialty share.

### Procedural Concentration and the Volume–Outcome Relationship

The Gini coefficient and HHI provide complementary views of concentration, and both supported the same conclusion. Moderate inequality across the full distribution (Gini `r sprintf("%.2f", v$urps_gini)`–`r sprintf("%.2f", v$gob_gini)`, with overlapping confidence intervals across the well-populated groups) coexisted with size-adjusted normalized HHI values near zero and effective provider counts in the hundreds. In other words, sling volume was distributed unevenly across many surgeons but was not dominated by a few, and the well-populated specialties did not differ meaningfully in concentration. Because the physician was the production unit, these values describe surgeon-level procedural concentration rather than market competition, and they should not be interpreted using antitrust thresholds.

These patterns matter because greater surgeon volume has been associated with better outcomes after midurethral sling surgery. Berger et al. reported a lower adjusted risk of reoperation for sling failure among higher-volume surgeons.^16^ A systematic review by Cartier et al. found that low-volume surgeons had higher odds of mesh revision and repeat incontinence procedures.^17^ Brennand and Quan observed lower revision odds above approximately 50 cases per year,^18^ and Holdø and Svenningsen found better objective cure rates with greater surgeon experience and annual volume.^19^ The observed fee-for-service Medicare medians of `r v$gob_med`–`r v$urps_med` services per year cannot be compared directly with all-payer thresholds such as approximately 50 operations per year, because the PUF omits commercially insured, Medicare Advantage, and younger patients and providers below the CMS suppression threshold; total surgeon experience is therefore higher than the observed Medicare median, and Medicare PUF volume is an incomplete proxy for it.

### Temporal Trends and the Evolving SUI Landscape

The repeated-measures model and turnover analysis clarify why the URPS share increased. Individual physician volume did not change over time, but the number of non-URPS surgeons fell while the URPS pool remained comparatively stable. URPS also contributed the largest inflow of newly observable surgeons, whereas urology had substantial turnover but a net decline. The market-share shift therefore reflects workforce composition within a shrinking procedure pool, not increasing individual volume or concentration.

This mechanism is consistent with prior work. Siegal et al. found that the post-2011 decline in sling placement was driven mainly by non-FPMRS providers, while FPMRS providers maintained their volumes.^12^ Lee et al. documented an approximately 50% national decline in incontinence surgery from 2004 through 2013.^13^ The 2020 decline in our study (RR `r sub(" .*", "", v$gee_covid)`) is consistent with pandemic-related deferral of elective surgery and remained robust when 2020 was excluded.

The SUI treatment landscape also continues to change. Urethral bulking agents are now positioned alongside slings as a first-line surgical option,^3,20^ and ongoing surveillance is needed to determine whether this shift affects specialty-specific sling volume.

### Limitations

This study has five limitations. First, CMS suppresses provider-level data for fewer than 11 beneficiaries, so the lowest-volume surgeons are not observed; suppression can bias the Gini coefficient and HHI in different directions, and the full-market concentration cannot be identified from the Public Use File, although a sensitivity analysis adding hypothetical suppressed providers left the low-concentration conclusion unchanged. Second, the PUF includes only fee-for-service Medicare Part B claims; younger patients and those covered by Medicare Advantage, Medicaid, commercial insurance, or no insurance are not represented.

Third, specialty classification depends on CMS provider type and the ABOG/ABU rosters. Physicians who completed URPS fellowship but were not board-certified may therefore be misclassified, and the `r v$class_reclass_urology` providers reclassified from an adjacent provider type to urology could be misassigned. Fourth, the market-share trend is estimated under plausible classification scenarios rather than point-identified. Certification follows the start of subspecialty practice, so the certification-gated scenario assigns some pre-certification practice years to General OB/GYN or urology, whereas fixed membership assigns physicians to URPS before they subspecialized. We present these as plausible scenarios, not formal bounds; point identification would require unavailable fellowship-completion dates.

Approximately 8% of URPS physicians could not be matched to a certification year, and urology-pathway URPS classification remained fixed because no urology subspecialty date was available; both features narrow the certification-gated estimate. Fifth, Tot_Srvcs may include bilateral or modifier-inflated services.

### Strengths

Strengths include a national sample of all publicly reported, nonsuppressed fee-for-service provider-service records; identification of urogynecologists through both certification pathways; complementary surgeon-level concentration measures with bootstrap confidence intervals, calculated both overall and by year; models that account for repeated observations within physicians; transparent handling of ambiguous and facility billers with sensitivity analyses; and a fully reproducible analytic pipeline.

## Conclusions

Among physician-years observable in fee-for-service Medicare, URPS physicians performed most reported sling services for stress urinary incontinence, and their share increased from 2013 through 2023. The increase was driven primarily by the OB/GYN pathway and declining observable participation among non-URPS clinicians, not by increasing surgeon-level concentration. Because the PUF excludes low-volume physician-years and Medicare Advantage claims and does not distinguish synthetic midurethral from fascial sling procedures, these results describe observable Medicare service patterns rather than the complete national sling market or clinical workforce. All-payer studies linking volume, outcomes, and geographic access are needed before drawing credentialing conclusions.

## References

1. Wu JM. Stress incontinence in women. N Engl J Med. 2021;384(25):2428-2436.
2. Ford AA, Rogerson L, Cody JD, Aluko P, Ogah JA. Mid-urethral sling operations for stress urinary incontinence in women. Cochrane Database Syst Rev. 2017;7(7):CD006375.
3. Kobashi KC, Vasavada S, Bloschichak A, et al. Updates to surgical treatment of female stress urinary incontinence (SUI): AUA/SUFU guideline (2023). J Urol. 2023;209(6):1091-1098.
4. Birkmeyer JD, Stukel TA, Siewers AE, Goodney PP, Wennberg DE, Lucas FL. Surgeon volume and operative mortality in the United States. N Engl J Med. 2003;349(22):2117-2127.
5. Reames BN, Ghaferi AA, Birkmeyer JD, Dimick JB. Hospital volume and operative mortality in the modern era. Ann Surg. 2014;260(2):244-251.
6. American Board of Medical Specialties. ABMS announces approval of new subspecialty certificate in female pelvic medicine and reconstructive surgery. 2011. Available at: https://www.abms.org.
7. Berger AA, Tan-Kim J, Menefee SA. The impact of the 2011 US Food and Drug Administration transvaginal mesh communication on utilization of synthetic mid-urethral sling procedures. Int Urogynecol J. 2021;32(8):2227-2231.
8. Stitzenberg KB, Sigurdson ER, Egleston BL, Starkey RB, Meropol NJ. Centralization of cancer surgery: implications for patient access to optimal care. J Clin Oncol. 2009;27(28):4671-4678.
9. James MB, Theofanides MC, Sui W, Onyeji I, Badalato GM, Chung DE. Sling procedures for the treatment of stress urinary incontinence: comparison of national practice patterns between urologists and gynecologists. J Urol. 2017;198(6):1386-1391.
10. Cantrell AB, Rothschild J, Durbin-Johnson B, Gonzalez R, Kurzrock EA. Surgical trends in the correction of female stress urinary incontinence in academic centers within the United States. Neurourol Urodyn. 2017;36(2):394-398.
11. Rogo-Gupta L, Litwin MS, Saigal CS, Anger JT; Urologic Diseases in America Project. Trends in the surgical management of stress urinary incontinence among female Medicare beneficiaries, 2002-2007. Urology. 2013;82(1):38-41.
12. Siegal AR, Huang Z, Gross MD, Mehraban-Far S, Weissbart SJ, Kim JM. Trends of mesh utilization for stress urinary incontinence before and after the 2011 Food and Drug Administration notification between FPMRS-certified and non-FPMRS-certified physicians: a statewide all-payer database analysis. Urology. 2021;150:151-157.
13. Lee UJ, Feinstein L, Ward JB, et al; Urologic Diseases in America Project. National trends in the surgical management of urinary incontinence among insured women, 2004 to 2013. J Urol. 2020;203(2):365-371.
14. Tabakin AL, Sawhney R, Daily AM, Winkler HA, Shalom DF, Tam J, Lee W. Case log trends of urogynecology and reconstructive pelvic surgery fellows: a comparison of urology- and gynecology-based fellowship programs. Neurourol Urodyn. 2024;43(8):1970-1976.
15. Wang CN, Su IW, Smith AL, Badalato GM, Chung DE. Current exposure to female pelvic medicine and reconstructive surgery faculty during urology residency. Neurourol Urodyn. 2023;42(7):1569-1573.
16. Berger AA, Tan-Kim J, Menefee SA. Surgeon volume and reoperation risk after midurethral sling surgery. Am J Obstet Gynecol. 2019;221(5):523.e1-523.e8.
17. Cartier S, Cerantola GM, Leung AA, Brennand E. The impact of surgeon operative volume on risk of reoperation within 5 years of mid-urethral sling: a systematic review. Int Urogynecol J. 2023;34(5):981-992.
18. Brennand EA, Quan H. Evaluation of the effect of surgeon's operative volume and specialty on likelihood of revision after mesh midurethral sling placement. Obstet Gynecol. 2019;133(6):1099-1108.
19. Holdø B, Svenningsen R. The impact from surgical experience on short- and long-term success rates after mid-urethral sling surgery. Int Urogynecol J. 2026 (epub ahead of print).
20. Gallo K, Weiner H, Mishra K. An update on surgical management for stress urinary incontinence. Curr Opin Obstet Gynecol. 2024;36(6):433-438.

---

## Tables

**Table 1.** Specialty distribution and market-share trends among clinicians billing CPT 57288, Medicare PUF 2013–2023, with URPS split by certification pathway. Δ share is the annual percentage-point change from ordinary least-squares regression on calendar year. Specialty-specific physician counts exceed `r format(v$analytic_physicians, big.mark=",")` because `r v$switchers` physicians changed groups across years. *MIGS estimates are descriptive because only `r v$mig_phys` physicians were identified.

```{r t1}
kt(tab$t1)
```

**Table 2.** Surgeon-level procedural concentration by specialty, based on aggregate provider volume. Raw HHI is not comparable across groups of different size; the normalized HHI (0–1) and effective number of providers are size-adjusted companions.

```{r t2}
kt(tab$t2)
```

**Table 3.** Adjusted rate ratios for annual reported services from a Poisson GEE clustered by NPI (OB/GYN-pathway URPS reference; calendar year centered at 2018; exchangeable correlation; robust standard errors). The first block gives specialty contrasts at mid-study and the 2020 effect; the second block gives each specialty's annual trend as a marginal contrast (the year term plus its specialty-by-year interaction).

```{r t3}
kt(tab$t3)
```

**Table 4.** All-pathway URPS and OB/GYN-based market-share trends under alternative classification scenarios. The scenarios differ in the estimated rate of increase: fixed membership gives the shallower slope because it counts physicians as URPS before certification, and certification-gated classification gives the steeper slope because it removes not-yet-certified physicians from the early URPS count. The 2023 levels converge. These are plausible scenarios rather than formal statistical bounds. All estimated trends are positive and statistically significant.

```{r t4}
kt(tab$t4)
```

---

## Figures

```{r fig1, results='asis'}
show_fig("figure_rate_per_100k.png")
```

**Figure 1.** Utilization of sling surgery for stress urinary incontinence (CPT 57288) in fee-for-service Medicare, 2013–2023. Bars show reported services; the line shows the denominator-adjusted rate per 100,000 female Part B fee-for-service beneficiaries. The rate declined less than the raw count (`r format(v$rate_pct_change, nsmall=1)`% versus `r format(v$count_pct_change, nsmall=1)`%) because the fee-for-service population contracted by `r format(v$den_pct_change, nsmall=1)`%; the linear trend in the rate was not significant (p = `r v$rate_slope_p`).

```{r fig2, results='asis'}
show_fig("figure_market_share.png")
```

**Figure 2.** Market share of sling operations for stress urinary incontinence (CPT 57288) by specialty, fee-for-service Medicare PUF 2013–2023. Stacked areas are the fixed-membership specialty shares; the two black lines are the URPS share under fixed-membership (solid) and certification-gated (dashed) classification. Fixed membership (all-pathway URPS share `r v$fixed_urps_s13`% to `r v$fixed_urps_s23`%; +`r sprintf("%.2f", v$fixed_urps_slope)` percentage points/year) and certification-gated classification (`r v$ct_urps_s13`% to `r v$ct_urps_s23`%; +`r sprintf("%.2f", v$ct_urps_slope)` percentage points/year) differ in the estimated rate of increase and converge by 2023.

```{r fig3, results='asis'}
show_fig("figure_classification_flow.png")
```

**Figure 3.** Provider-classification flow for CPT 57288. Billers are assigned to specialty groups by CMS provider type cross-referenced with the ABOG and ABU rosters. Billers with neither an OB/GYN nor urology CMS type and no ABOG match (`r v$class_reclass_urology` NPIs) are predominantly organizations and non-physician clinicians; organizational NPIs (NPPES entity type 2: ambulatory surgical centers, hospitals, laboratories; `r v$class_excluded_facility`) are excluded and the remaining individual clinicians (`r v$other_phys`) form an Other/uncertain group, rather than being counted as urologists. Specialty-specific counts exceed the analytic cohort because some physicians changed groups across years.

---

## Supplementary Information

Supplementary tables are generated by pipeline step 06 from the same frozen cache as the main results, so they remain consistent with the numbers above.

**Supplementary Table S1.** Annual procedural concentration by calendar year and specialty. Total procedures, observable surgeons, median annual volume, Gini coefficient, HHI, and top-decile, top-quintile, and bottom-half shares are reported for every year. Within-year concentration was low and stable across the study period.

```{r s1}
s <- read_supp("table_5_annual_concentration")
if (!is.null(s)) kt(s) else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S2.** Trend regressions for annual concentration and volume measures on calendar year, overall and by specialty (ordinary least squares).

```{r s2}
s <- read_supp("table_6_concentration_trends")
if (!is.null(s)) {
  s <- dplyr::mutate(s,
    `Slope / year` = round(as.numeric(`Slope / year`), 3),
    `R-squared`    = round(as.numeric(`R-squared`), 3),
    `p-value`      = fmt_pcol(`p-value`))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S3.** Specialty-specific market-share trends. Each specialty's annual share of observed procedures is regressed on calendar year separately (fixed-membership classification).

```{r s3}
s <- read_supp("table_12_specialty_share_trends")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Specialty       = specialty_group,
    `2013 share (%)`= share_start,
    `2023 share (%)`= share_end,
    `Slope (pp/yr)` = round(as.numeric(slope_pp_yr), 3),
    `p-value`       = fmt_pcol(p_value))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S4.** Adjusted rate ratios for annual sling volume from a Poisson GEE clustered by NPI, excluding calendar year 2020 (COVID-19 sensitivity analysis; URPS reference).

```{r s4}
s <- read_supp("table_8b_volume_gee_no2020")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Term          = gsub("specialty_group", "Specialty: ", term),
    `RR (95% CI)` = `RR (95% CI)`,
    `p-value`     = fmt_pcol(p_value))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S5.** Per-physician volume comparisons (one median value per physician), Kruskal-Wallis and pairwise Wilcoxon tests with Bonferroni correction.

```{r s5}
s <- read_supp("table_9_per_physician")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Test        = test,
    Statistic   = ifelse(is.na(suppressWarnings(as.numeric(statistic))), "n/a",
                         sprintf("%.2f", suppressWarnings(as.numeric(statistic)))),
    df          = df,
    `p-value`   = p_formatted)
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S6a.** Specialty distribution under alternative classification schemes (time-varying, modal, ever-URPS/MIGS, certification-gated).

```{r s6a}
s <- read_supp("table_10_classification_sensitivity")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Scheme       = scheme,
    Specialty    = specialty_group,
    `N providers`= n_providers,
    Procedures   = procedures,
    `% of all`   = sprintf("%.1f%%", as.numeric(pct_of_all)))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S6b.** All-pathway URPS and OB/GYN-based market-share trends under each classification scheme. The URPS increase remains positive and significant in every scheme.

```{r s6b}
s <- read_supp("table_10b_classification_trends")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Scheme               = scheme,
    `URPS slope (pp/yr)` = round(as.numeric(urps_slope_pp_yr), 3),
    `URPS p`             = fmt_pcol(urps_p),
    `Gyn slope (pp/yr)`  = round(as.numeric(gyn_slope_pp_yr), 3),
    `Gyn p`              = fmt_pcol(gyn_p))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S7.** Observable surgeons and URPS share by provider practice state (geography, secondary analysis). A state with no observable URPS surgeon may still have URPS surgeons who each fell below the CMS reporting threshold.

```{r s7}
s <- read_supp("table_13_geography_by_state")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    State                 = state,
    `Observable surgeons` = n_observable_surgeons,
    `URPS surgeons`       = n_urps,
    `URPS share`          = sprintf("%.1f%%", as.numeric(urps_share_pct)),
    `Observable URPS?`    = ifelse(as.logical(has_observable_urps), "Yes", "No"))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S8.** Suppression sensitivity. Hypothetical suppressed low-volume providers (25% and 50% of the observed count, each performing 1–10 services) were added to each group before recomputing the Gini coefficient and HHI. Concentration rose only modestly, so the low-concentration conclusion is robust to the unobserved low-volume tail.

```{r s8}
if (!is.null(tab$t_suppress)) kt(tab$t_suppress) else cat("_Not available._")
```

**Supplementary Table S9.** Annual observable participation transitions using a two-year washout. Newly observable surgeons were absent in both prior observable years; surgeons no longer observable were absent in both subsequent observable years. Newly observable and continuing counts are undefined for the first two years, and no-longer-observable counts are undefined for the last two years. These describe threshold crossings, not definitive entry into or exit from practice.

```{r s9}
if (!is.null(tab$t5)) kt(tab$t5) else cat("_Not available._")
```

Newly observable episodes by specialty: URPS, `r v$ent_urps`; urology, `r v$ent_uro`; other OB/GYN, `r v$ent_gob`; MIGS, `r v$ent_migs`.

**Supplementary Table S11.** Classification sensitivity for ambiguous and facility billers (reviewer concern). The primary analysis excludes facility NPIs and places remaining non-physician or unclassifiable clinicians in an "Other/uncertain" group. Assigning them to urology (the legacy approach) roughly doubles the apparent non-URPS urology share; excluding them entirely changes the shares little for the well-populated groups. The URPS increase is significant under all three.

```{r s11}
if (!is.null(tab$t_classif_sens)) kt(tab$t_classif_sens) else cat("_Not available._")
```

---

## Supplementary Figures

```{r sfig1, results='asis'}
show_fig("figure_2_volume_distribution.png")
```

**Supplementary Figure S1.** Annual sling service volume (CPT 57288) by specialty (violin and box plots; logarithmic scale). The minimum observable volume is 11 because of CMS cell suppression.

```{r sfig2, results='asis'}
show_fig("figure_3_lorenz_curve.png")
```

**Supplementary Figure S2.** Lorenz curves of procedural concentration by specialty. Greater distance from the diagonal indicates greater concentration.

```{r sfig3, results='asis'}
show_fig("figure_4_concentration_trends.png")
```

**Supplementary Figure S3.** Annual within-year concentration by specialty (Gini, HHI, top-20% share, bottom-50% share). Within-year concentration was low and stable. MIGS was excluded because too few surgeons were observed each year for stable estimates.

```{r sfig4, results='asis'}
show_fig("figure_5_supply_trends.png")
```

**Supplementary Figure S4.** Observable surgeons and total reported services by year and specialty. The decline in the observable surgeon pool was concentrated in urology and General OB/GYN.
