---
title: "Specialty Distribution and Surgeon-Level Concentration of Sling Surgery for Stress Urinary Incontinence in Fee-for-Service Medicare, 2013–2023"
author: "Tyler Muffly, MD"
output:
  word_document:
    # Urogynecology (LWW) submissions are double-spaced with continuous line
    # numbering. Pandoc's default reference doc is single-spaced and unnumbered,
    # so the template carries both. The path is relative to this Rmd's own
    # folder (output/), not to knit_root_dir: pandoc resolves resource paths
    # from where the Rmd lives, which is the same quirk show_fig() works around
    # with absolute paths below.
    reference_docx: ../templates/urogynecology_reference.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
suppressWarnings(suppressMessages({
  library(dplyr); library(knitr); library(readr)
}))
# Supplementary tables and figures are produced by pipeline steps 06 and 07
# from the same cache, so they stay consistent with the inline numbers. Both
# helpers degrade gracefully if an artifact is missing.
# Paths are relative to knit_root_dir (the project root), which is where chunk
# *code* runs. Guards live inside chunk code, not in the `eval` option, because
# knitr evaluates chunk options relative to the Rmd's own folder (output/).
fig_dir    <- "output/figures"
tables_dir <- "output/tables"
# Emit a raw-markdown image with an ABSOLUTE path. Pandoc resolves absolute
# paths regardless of its working directory, so the PNG always embeds into the
# docx. (include_graphics() records a root-relative path that pandoc, run from
# the Rmd's folder, cannot find, so it silently drops the image.)
show_fig <- function(name, width = "6.5in") {
  p <- normalizePath(file.path(fig_dir, name), mustWork = FALSE)
  if (file.exists(p)) cat(sprintf("\n![](%s){width=%s}\n\n", p, width))
  else cat(sprintf("\n_Figure not found (%s); run pipeline step 07._\n\n", name))
}
read_supp <- function(name) {
  p <- file.path(tables_dir, paste0(name, ".csv"))
  if (!file.exists(p)) return(NULL)
  suppressWarnings(readr::read_csv(p, show_col_types = FALSE))
}
fmt_pcol <- function(x) {
  x <- suppressWarnings(as.numeric(x))
  ifelse(is.na(x), "n/a", ifelse(x < 0.001, "<0.001", sprintf("%.3f", x)))
}
# Every number below is computed from the pipeline cache, not hardcoded, so the
# prose can never drift from the analysis. The path defaults to the pipeline
# cache and can be overridden with the PUF_CLASSIFIED environment variable.
source("R/compute_manuscript_values.R")
cache_dir <- tryCatch(config::get("cache_dir"), error = function(e) "data/cache")
puf_path  <- Sys.getenv("PUF_CLASSIFIED", file.path(cache_dir, "puf_classified.rds"))
mv  <- compute_manuscript_values(puf_path)
v   <- mv$v
tab <- mv$tab
kt  <- function(df, align = NULL) knitr::kable(df, align = align, format = "pipe")
# composite inline helpers
iqr  <- function(m, lo, hi) sprintf("%g (%g–%g)", m, lo, hi)
ci   <- function(s, lo, hi, p) sprintf("%+.2f percentage points/year; 95%% CI %.2f to %.2f; p %s",
                                       s, lo, hi, ifelse(p == "<0.001", "< 0.001", paste0("= ", p)))
```

**Author:** Tyler Muffly, MD^1^

^1^ *[Affiliation to be completed: division, department, institution, city, state]*

**Corresponding author:** Tyler Muffly, MD. *[mailing address, telephone, and email to be completed]*

**Funding:** None.

**Conflicts of interest:** *[to be completed; state "The author reports no conflicts of interest" if applicable]*

**Acknowledgments:** *[to be completed]*

**Data and code availability:** The analytic code and non-restricted derived outputs, together with this manuscript source, are available from the corresponding author on reasonable request. The CMS Medicare Physician and Other Practitioners Public Use File and the CMS Program Statistics Original Medicare Enrollment tables are publicly available from data.cms.gov. The ABOG and ABU board-certification rosters used for specialty linkage are subject to source-specific permissions and may not be redistributable; provenance and derived crosswalk documentation are available with the analytic code.

**Manuscript type:** Original research.

**Running head:** URPS Share of Medicare Sling Surgery

**Precis:** Urogynecologists performed most observable sling procedures in fee-for-service Medicare from 2013 through 2023, and their share rose without a corresponding increase in surgeon-level concentration.

<!-- Word counts are document properties rather than analysis results, so unlike
     every number below they are not computed from the cache. Re-measure them on
     the final rendered docx immediately before submission if the text changes. -->
**Word count:** Abstract 302 words; main text (Introduction through Conclusions) 4,078 words. Tables: 4. Figures: 3. References: 20.

---

## Abstract

**Objectives:** To describe the specialty distribution, annual clinician volume, and surgeon-level concentration of sling surgery for stress urinary incontinence (CPT 57288) in fee-for-service Medicare from 2013 through 2023.

**Methods:** We analyzed all 11 annual CMS Medicare Physician and Other Practitioners Public Use Files. Physicians billing CPT 57288 were classified by annual CMS provider type and ABOG and ABU roster linkage into all-pathway URPS (OB/GYN or urology certification), non-URPS urology, other non-URPS OB/GYN, or MIGS; organizational and unclassifiable billers were excluded from the primary cohort and examined separately. The primary market-share estimand was the all-pathway URPS share under fixed-membership classification. We modeled utilization with female Part B fee-for-service enrollment as an offset, annual volume with a Poisson generalized estimating equation clustered by NPI, and URPS share with a quasibinomial model. Annual Gini coefficients measured concentration.

**Results:** Among `r format(v$analytic_physicians, big.mark=",")` physicians, `r format(v$analytic_procs, big.mark=",")` sling services were observable. URPS physicians, identified through either certification pathway, performed the majority (`r v$urps_combined_pct`%; OB/GYN pathway `r v$urps_pct`%, urology pathway `r v$urpsuro_pct`%); non-URPS urology performed `r v$uro_pct`% and other non-URPS OB/GYN `r v$gob_pct`%. Reported services declined `r format(abs(v$count_pct_change), nsmall=1)`%; after enrollment adjustment, the estimated utilization rate declined `r sub("^-","",v$rate_offset_pctyr)`% per year (rate ratio `r v$rate_offset_rr`, 95% CI `r v$rate_offset_ci`; p = `r v$rate_offset_p`). Under fixed-membership classification, the all-pathway URPS share increased from `r v$fixed_urps_s13`% in 2013 to `r v$fixed_urps_s23`% in 2023 (+`r sprintf("%.2f", v$fixed_urps_slope)` percentage points/year; p `r ifelse(v$fixed_urps_p=="<0.001","< 0.001",paste("=",v$fixed_urps_p))`); the increase remained significant under certification-gated classification. OB/GYN-pathway URPS had the highest adjusted annual volume, and within-year surgeon concentration remained stable.

**Conclusions:** In observable fee-for-service Medicare, URPS physicians identified through either certification pathway performed most sling services, and their share increased without a corresponding increase in surgeon concentration. These findings describe claims-visible service patterns, not the complete national surgical workforce; all-payer outcome and access studies are needed before drawing credentialing implications.

**Keywords:** stress urinary incontinence; sling surgery; urogynecology; Medicare; surgical workforce.

---

## Simply Stated

Sling surgery treats stress urinary incontinence, which causes urine leakage with coughing, laughing, or exercise. We studied traditional Medicare billing records from 2013 through 2023 to learn which doctors performed these operations and how the work was shared. Urogynecologists performed most of the reported surgeries, and their share increased over time. The total number of reported surgeries fell. Some of that decline occurred because fewer people were enrolled in traditional Medicare, although a smaller true decline may remain. The work stayed spread across many surgeons rather than becoming centered in a small group. These data cannot show which specialty has the best outcomes or who should be allowed to perform the operation.

## Why This Matters

National data on who performs sling surgery are limited. This study improves classification by identifying urogynecologists through both the obstetrics-gynecology and urology certification pathways, avoiding the misclassification of urology-trained urogynecologists as general urologists. It also pairs claims with fee-for-service enrollment, showing that raw service counts overstate the decline in utilization. The main finding is a growing URPS share without increasing within-year surgeon concentration. These results support workforce surveillance, but claims data alone cannot establish quality, access, or credentialing standards.

---

## Introduction

Sling surgery is a standard surgical treatment for stress urinary incontinence (SUI) when conservative treatment is insufficient.^1,2^ CPT 57288 describes a "sling operation for stress incontinence (eg, fascia or synthetic)." The code therefore includes both autologous fascial and synthetic sling procedures and does not distinguish sling material or retropubic, transobturator, single-incision, and pubovaginal techniques. Despite the procedure's widespread use, little is known about which specialties perform it in the Medicare population or how case volume is distributed among surgeons.

These questions are relevant to surgical quality, training, and workforce planning. Across surgical fields, greater surgeon and hospital volume is generally associated with better outcomes,^4,5^ although procedure-specific volume thresholds for sling surgery remain uncertain. Changes in the specialty mix may also affect training needs and patient access.

Three developments make this analysis timely. First, Female Pelvic Medicine and Reconstructive Surgery, approved as a subspecialty by the American Board of Medical Specialties in 2011 with the first certificates issued in 2013 and renamed Urogynecology and Reconstructive Pelvic Surgery (URPS) in 2024, has expanded rapidly.^6^ Certification is available through both the American Board of Obstetrics and Gynecology (ABOG) and the American Board of Urology (ABU). Second, US Food and Drug Administration actions involving urogynecologic mesh changed public perceptions and may have affected sling use.^7^ The 2019 order concerned transvaginal mesh for pelvic organ prolapse; it did not remove synthetic midurethral slings for SUI from the market. Third, the distribution of procedures across clinicians has become increasingly relevant to quality improvement and resource planning.^8^

We therefore examined specialty distribution, annual clinician volume, and surgeon-level concentration among clinicians reporting CPT 57288 services in fee-for-service Medicare from 2013 through 2023. We also evaluated changes in each specialty's share of observable services.

## Materials and Methods

### Study Design and Data Source

We conducted a repeated cross-sectional analysis of the CMS Medicare Physician and Other Practitioners Public Use File (PUF) from 2013 through 2023. The PUF contains publicly reportable fee-for-service Medicare Part B claims aggregated to the provider-service level. Each row represents a unique combination of National Provider Identifier (NPI), Healthcare Common Procedure Coding System code, place of service, and calendar year. Because the data are public and contain no protected health information, institutional review board approval was not required.

Before analysis, we verified all 11 annual PUF releases against the CMS file specifications and confirmed that each expected year was complete. We report the study in accordance with the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) guideline for cross-sectional studies.

### Specialty Classification

Each billing NPI was assigned to one of five analyzed physician groups (URPS through the OB/GYN pathway, URPS through the urology pathway, Minimally Invasive Gynecologic Surgery (MIGS), other non-URPS OB/GYN, or non-URPS urology) or to a residual Other/uncertain category that we set aside from the primary cohort. We first used the annual CMS rendering-provider type, which may vary by year, to identify OB/GYN, urology, and other clinicians. Among OB/GYN physicians, the ABOG registry identified diplomates certified in Female Pelvic Medicine and Reconstructive Surgery as URPS and those with a MIG designation as MIGS. Records were linked primarily by exact NPI and, when necessary, by full name and practice state. Remaining OB/GYN physicians were classified as other non-URPS OB/GYN.

Among `r v$class_reclass_urology` NPIs with neither an OB/GYN nor urology CMS provider type and no ABOG match, most represented organizations or nonphysician clinicians rather than urologists, and none matched the ABU roster. We excluded `r v$class_excluded_facility` organizational NPIs (NPPES entity type 2, flagged by a rendering-provider entity code of "O") because they represented billing entities, such as ambulatory surgery centers, hospitals, laboratories, or groups, rather than individual clinicians. The remaining `r v$other_phys` individual nonphysician or unclassifiable billers (`r v$other_pct`% of observable services) could not be confirmed as either specialty; we excluded them from the primary physician cohort and retained them as a separate Other/uncertain group only in sensitivity analyses. Figure 3 shows the full classification flow. Supplementary Table S11 compares this primary approach with retaining these billers as a separate group and with the legacy approach of assigning them to urology; the legacy approach approximately doubled the apparent non-URPS urology share.

We then cross-referenced the ABU roster of `r nrow(readr::read_csv("data/abu_urology/abu_urps_npi_LATEST.csv", show_col_types=FALSE))` urology-pathway urogynecologists. Of these, `r v$class_abu_pathway` appeared in the analytic cohort and were classified as urology-pathway URPS; all remaining urology physicians were classified as non-URPS urology. Without this linkage, fellowship-trained urology-pathway urogynecologists would appear as general urologists in the PUF. We therefore report all-pathway URPS separately from the OB/GYN-based share, because all-pathway URPS includes urology-trained subspecialists who are not gynecologists.

Because 45% of URPS physicians in the cohort were certified after 2013, the treatment of pre-certification years could affect the estimated market-share trend. We therefore used two plausible scenarios with opposing assumptions. Fixed membership assigned each physician's eventual subspecialty to every study year and therefore tended to produce a shallower slope by counting physicians as URPS before certification.

Certification-gated classification assigned URPS or MIGS status only beginning in the ABOG subspecialty certification year (sub1startdate, distinct from the initial OB/GYN board date). Earlier years were classified using the annual CMS provider type. Because certification generally follows the start of subspecialty practice, this scenario may misclassify genuine pre-certification subspecialty practice and therefore tends to produce a steeper slope. We treat the two approaches as plausible scenarios rather than formal bounds.

Modal and ever-URPS/MIGS classifications were additional sensitivity analyses. Urology-pathway URPS status remained fixed because the ABU roster did not include a subspecialty certification date. For pooled distribution, concentration, and volume analyses (Tables 1–3), each physician was assigned to the eventual combined specialty group.

### CMS Data Suppression

CMS suppresses provider-service records when a clinician treats fewer than 11 Medicare beneficiaries for a given code in a calendar year. Every observable physician-year in this study therefore represents at least 11 beneficiaries with CPT 57288. Clinicians below that threshold are not visible, so the number and volume of low-volume surgeons cannot be recovered from the PUF. Because suppression may affect the Gini coefficient and HHI differently, full-market concentration is not identified. We therefore report bootstrap confidence intervals and a sensitivity analysis that adds hypothetical suppressed low-volume providers before recalculating concentration.

### Outcome Measures and Statistical Analysis

The primary clinician-level outcome was annual reported CPT 57288 services per provider (Tot_Srvcs). To describe utilization, we calculated observable services per 100,000 female Part B fee-for-service beneficiaries using CMS Program Statistics Original Medicare enrollment counts (table MDCR ENROLL AB 11, Sex = Female, Part B). We modeled annual service counts with log female fee-for-service enrollment as an offset and included a 2020 indicator. Because the PUF does not report beneficiary age for the numerator, this rate is crude and not age standardized.

We assessed surgeon-level concentration with complementary measures. The Gini coefficient summarizes inequality across the surgeon-volume distribution, and its 95% confidence interval was estimated by bootstrapping providers. The Herfindahl–Hirschman Index (HHI) is more sensitive to the largest-volume surgeons, but its minimum depends on the number of providers. We therefore report raw HHI descriptively and also present normalized HHI and the effective number of providers. Within-year concentration was primary; pooled multi-year measures were secondary because they also reflect the number of years each physician remained observable. The physician, rather than the hospital or health system, was the production unit, so these measures should not be interpreted using antitrust thresholds.

Because physicians could contribute up to 11 annual observations, we modeled annual reported services with a Poisson generalized estimating equation clustered by NPI, using an exchangeable working correlation and robust standard errors. The model was conditional on the physician-year being observable above the CMS threshold. Fixed effects included specialty (OB/GYN-pathway URPS reference), calendar year centered at 2018, specialty-by-year interactions, and a 2020 indicator. Specialty main effects therefore represent adjusted rate ratios at 2018, and specialty-specific annual trends were reported as marginal contrasts. A negative-binomial mixed model with an NPI random intercept served as a sensitivity analysis. We also repeated the analysis after excluding 2020 and after restricting the cohort to clinicians observable in at least two years.

As a secondary analysis with one independent observation per physician, we compared each physician's median annual volume across specialties using the Kruskal-Wallis test followed by Bonferroni-adjusted pairwise Wilcoxon tests.

For the primary market-share analysis, we modeled OB/GYN-pathway URPS services as a proportion of all annual observable services using a quasibinomial generalized linear model with calendar year centered at 2018. We report the annual odds ratio and the marginal change in percentage points per year. Separate ordinary least-squares regressions of annual specialty shares are presented descriptively. MIGS was not emphasized inferentially because only `r v$mig_phys` physicians were identified.

To describe changes in observable participation, we used a two-year washout. A surgeon was newly observable when present in a year but absent from both preceding observable years, continuing when present and not newly observable, and no longer observable when absent from both subsequent observable years. We summarized their annual service share, median volume, and specialty distribution. Because CMS suppression can move clinicians above or below the reporting threshold, these categories describe observation episodes rather than true entry into or exit from clinical practice. A physician could contribute more than one newly observable episode after an intervening gap.

In a secondary geographic analysis, we summarized observable surgeons and URPS share by practice state and identified states with no observable URPS surgeon performing at least 11 Medicare slings in any study year. We did not calculate state-level population rates because state-by-sex fee-for-service enrollment denominators were outside the scope of this analysis.

Analyses were performed in R version 4.4, with package versions locked through renv. The complete analytic pipeline is available from the corresponding author on reasonable request.

## Results

### Cohort and Annual Volume

Across the full 11-year window, `r format(v$full_physicians, big.mark=",")` physicians billed CPT 57288, accounting for `r format(v$full_services, big.mark=",")` observed services in `r format(v$full_pyears, big.mark=",")` physician-year observations. Participation was intermittent: `r v$persist_1` physicians (`r v$persist_1_pct`%) appeared in one year, `r v$persist_2_5` (`r v$persist_2_5_pct`%) in 2–5 years, `r v$persist_6_10` (`r v$persist_6_10_pct`%) in 6–10 years, and `r v$persist_11` (`r v$persist_11_pct`%) in all 11 years. Thus, only a small proportion maintained an observable Medicare sling practice throughout the study window.

The observed count of reported services fell from `r format(v$procs_2013, big.mark=",")` in 2013 to `r format(v$procs_2023, big.mark=",")` in 2023 (`r format(v$count_pct_change, nsmall=1)`%). Over the same period the female Part B fee-for-service population contracted from `r v$den_2013_m` million to `r v$den_2023_m` million beneficiaries (`r format(v$den_pct_change, nsmall=1)`%). In a Poisson model of the annual service count with log fee-for-service enrollment as an offset and a 2020 indicator, the observable utilization rate declined `r v$rate_offset_pctyr`% per year (rate ratio `r v$rate_offset_rr`, 95% CI `r v$rate_offset_ci`; about `r v$rate_offset_decade`% over the decade), but with only 11 annual observations and the pandemic interruption this trend was not statistically robust (dispersion-adjusted p = `r v$rate_offset_p`). The rate reached a low of `r v$rate_2020` per 100,000 in 2020, consistent with pandemic-related deferral of elective surgery, then partially recovered. The number of observable surgeons also fell, from `r v$surg_2013` to `r v$surg_2023`. A clinically meaningful decline in observable utilization thus remains possible, but the raw service-count drop overstates it because much of that drop reflects the shrinking fee-for-service denominator; because CMS suppresses low-volume physician-years, this is an observable-services rate, not the total national fee-for-service rate.

### Specialty Distribution and Trends

URPS physicians, identified through either certification pathway, performed the majority of observable services (`r v$urps_combined_pct`%; Table 1). Separating the two pathways, OB/GYN-pathway URPS (ABOG-certified) was the largest single group: `r v$urps_phys` physicians, `r format(v$urps_procs, big.mark=",")` services (`r v$urps_pct`%), median annual volume `r v$urps_med` (p25–p75, `r v$urps_p25`–`r v$urps_p75`). Urology-pathway URPS (ABU roster) added `r v$urpsuro_phys` physicians and `r format(v$urpsuro_procs, big.mark=",")` services (`r v$urpsuro_pct`%; median `r v$urpsuro_med`). Non-URPS urology included `r v$uro_phys` physicians and `r format(v$uro_procs, big.mark=",")` services (`r v$uro_pct`%; median `r v$uro_med`), other non-URPS OB/GYN `r v$gob_phys` physicians and `r format(v$gob_procs, big.mark=",")` services (`r v$gob_pct`%; median `r v$gob_med`), and MIGS `r v$mig_phys` physicians (`r v$mig_pct`%; median `r v$mig_med`). Organizational and unclassifiable billers were excluded from this primary physician cohort; including them changed the major specialty shares only slightly (Supplementary Table S11).

Because `r v$switchers` physicians changed specialty groups across years, specialty-specific counts exceed the unique cohort total. The urology-pathway URPS group (`r v$abu_moved_phys` physicians, `r format(v$abu_moved_procs, big.mark=",")` services) would otherwise be classified as general urology, understating the subspecialty and overstating non-URPS urology.

Annual market shares moved in different directions (Table 1). URPS increased from `r v$urps_trend$s13`% in 2013 to `r v$urps_trend$s23`% in 2023 (`r ci(v$urps_trend$slope, v$urps_trend$lo, v$urps_trend$hi, v$urps_trend$p)`). Urology decreased from `r v$uro_trend$s13`% to `r v$uro_trend$s23`% (`r ci(v$uro_trend$slope, v$uro_trend$lo, v$uro_trend$hi, v$uro_trend$p)`), and General OB/GYN decreased from `r v$gob_trend$s13`% to `r v$gob_trend$s23`% (`r ci(v$gob_trend$slope, v$gob_trend$lo, v$gob_trend$hi, v$gob_trend$p)`). MIGS increased descriptively from `r v$mig_trend$s13`% to `r v$mig_trend$s23`% (+`r sprintf("%.2f", v$mig_trend$slope)` percentage points/year), but this estimate is based on only `r v$mig_phys` physicians. In the quasibinomial model of URPS services out of all annual services, which respects the compositional structure, the URPS share rose from a fitted `r v$share_binom_2013`% in 2013 to `r v$share_binom_2023`% in 2023 (odds ratio `r v$share_binom_or` per year; `r v$share_binom_pp` percentage points per year; p `r ifelse(v$share_binom_p=="<0.001","< 0.001",paste("=",v$share_binom_p))`), consistent with the descriptive ordinary least-squares estimate. Because total service counts declined, these changes represent redistribution among specialties rather than growth in services.

The increase in all-pathway URPS share remained significant under every classification scenario (Table 4). Fixed membership, which counts physicians as URPS before certification, produced the shallower change from `r v$fixed_urps_s13`% in 2013 to `r v$fixed_urps_s23`% in 2023 (+`r sprintf("%.2f", v$fixed_urps_slope)` percentage points/year). Certification-gated classification produced the steeper change from `r v$ct_urps_s13`% to `r v$ct_urps_s23`% (+`r sprintf("%.2f", v$ct_urps_slope)` percentage points/year). The estimates converged by the end of the study period. These are prespecified scenarios addressing uncertain pre-certification practice, not statistical confidence bounds.

In the Poisson GEE clustered by NPI (reference = OB/GYN-pathway URPS; calendar year centered at 2018), adjusted annual volume at mid-study was lower than OB/GYN-pathway URPS for every other group, including urology-pathway URPS (RR `r v$gee_urpsuro`), non-URPS urology (RR `r v$gee_urology`), other non-URPS OB/GYN (RR `r v$gee_gob`), and MIGS (RR `r v$gee_migs`) (Table 3). Annual volume was approximately 12% lower in 2020 (RR `r v$gee_covid`; p `r ifelse(v$gee_covid_p=="<0.001","< 0.001",paste("=",v$gee_covid_p))`). The specialty-by-year interaction showed that per-physician volume was essentially flat over time for the well-populated groups (OB/GYN-pathway URPS RR `r v$slope_urps`, p = `r v$slope_urps_p`; urology-pathway URPS RR `r v$slope_urpsuro`, p = `r v$slope_urpsuro_p`; non-URPS urology RR `r v$slope_uro`, p = `r v$slope_uro_p`; other OB/GYN RR `r v$slope_gob`, p = `r v$slope_gob_p`), increasing only for the 10-physician MIGS group (RR `r v$slope_migs`; p = `r v$slope_migs_p`) (Table 3). A negative-binomial mixed model with a random intercept per NPI reproduced the direction but with attenuated magnitude (urology RR `r if (v$has_nb) v$nb_urology else "NA"`, General OB/GYN RR `r if (v$has_nb) v$nb_gob else "NA"`, 2020 RR `r if (v$has_nb) v$nb_covid else "NA"`); the attenuation is expected because the GEE estimates a population-averaged (marginal) rate ratio whereas the mixed model estimates a physician-conditional one. The one-observation-per-physician secondary analysis agreed (Kruskal-Wallis H = `r v$kw_H`, df = `r v$kw_df`, p `r ifelse(v$kw_p=="<0.001","< 0.001",paste("=",v$kw_p))`); URPS volume exceeded urology and other OB/GYN in pairwise comparisons (p < 0.001), with no significant URPS-MIGS difference. Because these models condition on the physician-year being observable above the CMS reporting threshold, the estimand is annual service volume among observable physician-years, not full-workforce practice volume; if non-URPS physicians more often fall below the threshold, the URPS volume advantage may be overstated.

### Surgeon Volume and Concentration

Within-year concentration was the primary concentration analysis. The annual Gini coefficient remained stable, ranging from `r v$annual_gini_lo` to `r v$annual_gini_hi`, with no temporal trend (p = `r v$annual_gini_p`); the top 20% of observable surgeons performed approximately `r v$annual_top20`% of annual services. No specialty-specific annual Gini changed significantly. Thus, the contraction in the observable surgeon pool was not accompanied by increasing concentration among a smaller group of high-volume surgeons.

Pooled multi-year concentration was moderate: the Gini coefficient was `r v$urps_gini_ci` for OB/GYN-pathway URPS, `r v$uro_gini_ci` for non-URPS urology, and `r v$gob_gini_ci` for other non-URPS OB/GYN (Table 2). Bootstrap comparisons showed no difference between URPS and urology (difference `r v$ginidiff_urps_uro$txt`), whereas URPS was modestly less concentrated than other non-URPS OB/GYN (`r v$ginidiff_urps_gob$txt`, confidence interval excluding zero). Normalized HHI values were near zero. The effective provider count represented `r v$urps_effpct`% of OB/GYN-pathway URPS physicians, `r v$uro_effpct`% of non-URPS urologists, and `r v$gob_effpct`% of other non-URPS OB/GYN physicians, indicating broad distribution with somewhat greater inequality in the smaller groups. A sensitivity analysis that added hypothetical suppressed low-volume providers increased the Gini, particularly under the 50% scenario (Supplementary Table S8); it therefore supports the narrow conclusion that observable services were not dominated by a handful of surgeons, but not a precise estimate of full-market concentration. MIGS estimates were considered exploratory because only `r v$mig_phys` physicians were identified.

### Observable Participation Over Time

Using the two-year washout (Supplementary Table S9), `r v$ent_min`–`r v$ent_max` surgeons became newly observable each year and performed `r v$entpct_min`%–`r v$entpct_max`% of annual volume. Their median volume was low, at approximately `r v$ent_med` services, consistent with surgeons appearing just above the CMS suppression threshold. Newly observable counts fell to `r v$ent_2020` surgeons (`r v$entpct_2020`% of volume) in 2020, then rebounded to `r v$ent_2022` (`r v$entpct_2022`%) in 2022; the 2022 rebound may partly reflect re-observation of surgeons who fell below the threshold during the pandemic rather than genuinely new surgeons. Continuing surgeons (`r v$cont_min`–`r v$cont_max` per year) performed most annual services, while surgeons no longer observable ranged from `r v$exit_min` to `r v$exit_max` per year.

Across the study period, URPS accounted for the most newly observable episodes (`r v$ent_urps`), followed by other non-URPS OB/GYN (`r v$ent_gob`), non-URPS urology (`r v$ent_uro`), and MIGS (`r v$ent_migs`); these are episode counts, not unique physicians. The rising URPS share coincided with the largest inflow of newly observable surgeons together with a relatively stable continuing-URPS base (URPS surgeon count `r v$urps_surg_2013` to `r v$urps_surg_2023`; per-physician volume unchanged). Urology showed substantial turnover with net decline, and General OB/GYN declined in both observable surgeons and market share.

Because total service counts declined, the growth in URPS share represents a redistribution of a shrinking service pool toward URPS, not new procedure growth. These transitions describe crossings of the CMS reporting threshold, not definitive entry into or exit from practice.

Observable surgeons and the URPS share by practice state are reported as an exploratory secondary analysis in Supplementary Table S7. Because state-by-sex fee-for-service denominators were not available and CMS suppression removes low-volume providers, we do not draw access conclusions from these counts; population-based geographic analysis is reserved for a separate access-focused study.

## Discussion

This analysis yielded three principal findings. First, URPS physicians, identified through either certification pathway, performed the majority of observable sling services (`r v$urps_combined_pct`%; OB/GYN pathway `r v$urps_pct`%, urology pathway `r v$urpsuro_pct`%), and OB/GYN-pathway URPS had the highest adjusted annual volume. Second, under fixed-membership classification the all-pathway URPS share increased substantially, from `r v$fixed_urps_s13`% to `r v$fixed_urps_s23`%, driven primarily by the OB/GYN pathway, while non-URPS participation declined. Third, within-year surgeon concentration remained stable despite a smaller observable surgeon pool. These findings describe a shift in claims-visible fee-for-service practice rather than increasing individual volume or concentration. They should be interpreted as practice patterns, not direct evidence of entry into or exit from the clinical workforce.

### Specialty Distribution in Context

The two URPS pathways together performed `r v$urps_combined_pct`% of observable fee-for-service Medicare sling services: `r v$urps_pct`% through the OB/GYN pathway and `r v$urpsuro_pct`% through the urology pathway. Non-URPS urologists performed `r v$uro_pct`%. Because all-pathway URPS includes urology-trained subspecialists, it is not equivalent to a gynecologist share. The estimated non-URPS urology share was sensitive to classification: assigning facility and nonphysician billers to urology approximately doubled it (Supplementary Table S11), illustrating how a simple provider-type approach can inflate urology estimates. Comparisons with earlier studies require caution because datasets, specialty definitions, and care settings differed. James et al. reported that gynecologists performed 74.2% of sling procedures in ACS-NSQIP data from 2006 through 2013;^9^ Cantrell et al. found that urogynecologists performed 54% of stress-incontinence procedures at academic centers from 2009 through 2014;^10^ and Rogo-Gupta et al. found that urologists performed most Medicare sling operations from 2002 through 2007.^11^ The most defensible contemporary conclusion is that OB/GYN-pathway URPS was the largest single fee-for-service Medicare group performing CPT 57288.

Including both certification pathways was essential. Linking the ABU roster reclassified `r v$abu_moved_phys` urology-typed physicians, representing `r v$urpsuro_pct`% of all observable services, as urology-pathway URPS. This pathway-neutral definition reflects differences in operative case volume between urology- and gynecology-based URPS fellowships^14^ and variable URPS exposure during urology residency.^15^ Classifying these subspecialists as general urologists would understate the URPS share and overstate non-URPS urology.

### Procedural Concentration and the Volume–Outcome Relationship

The concentration measures offered complementary information. Within-year Gini coefficients were low and stable, while pooled multi-year Gini coefficients showed moderate inequality because they also reflected the number of years each surgeon remained observable. URPS and non-URPS urology had similar pooled concentration, whereas other non-URPS OB/GYN was modestly more concentrated. Normalized HHI values remained near zero, and effective provider counts were large. In practical terms, observable sling volume was spread across many surgeons rather than dominated by a few. Because physicians were the production unit, these measures describe surgeon-level procedural distribution, not hospital or health-system market competition.

This distribution is relevant because higher surgeon volume has been associated with better outcomes after midurethral sling surgery. Berger et al. reported lower adjusted reoperation risk among higher-volume surgeons,^16^ and a systematic review found greater risks of mesh revision and repeat incontinence procedures among low-volume surgeons.^17^ Other studies have reported lower revision odds above approximately 50 cases per year^18^ and better objective cure rates with greater experience and annual volume.^19^ The fee-for-service Medicare medians of `r v$gob_med`–`r v$urps_med` observable services per year cannot be compared directly with all-payer thresholds. The PUF excludes younger patients, Medicare Advantage, commercial insurance, and physician-years below the suppression threshold; total surgeon volume may therefore exceed the observed Medicare count, but cannot be inferred precisely.

### Temporal Trends and the Evolving SUI Landscape

The repeated-measures and participation analyses help explain the changing specialty shares. Annual volume among observable physicians was stable, but non-URPS participation declined while the URPS pool remained comparatively stable. URPS also accounted for the largest number of newly observable episodes. The shift therefore reflects a change in the specialty composition of a contracting observable service pool, not increasing individual volume or concentration.

This pattern is consistent with prior work. Siegal et al. found that the post-2011 decline in sling placement was driven mainly by non-FPMRS providers, whereas FPMRS providers maintained their volume.^12^ Lee et al. documented an approximately 50% national decline in incontinence surgery from 2004 through 2013.^13^ In the present study, the 2020 decrease (RR `r sub(" .*", "", v$gee_covid)`) was consistent with pandemic-related deferral of elective surgery, and the principal specialty comparisons were similar when 2020 was excluded.

The SUI treatment landscape is also evolving. Urethral bulking is increasingly used as a less-invasive procedural alternative for selected patients, although efficacy is generally lower and repeat treatment may be required,^3,20^ and continued surveillance will be needed to determine whether changing treatment preferences alter specialty-specific sling volume.

### Limitations

This study has several limitations. First, CMS suppresses records for clinicians treating fewer than 11 beneficiaries, so the lowest-volume physician-years are not observed. Full-market concentration cannot be identified, and the sensitivity analysis supports only the narrower conclusion that observable services were not dominated by a handful of surgeons. Second, the PUF includes only fee-for-service Medicare Part B claims. It excludes Medicare Advantage, commercial insurance, Medicaid, uninsured care, and most younger patients.

Third, specialty classification depends on annual CMS provider type and ABOG and ABU roster linkage. Fellowship-trained physicians who were not board certified may be misclassified, and the Other/uncertain group contains nonphysician and incompletely classified individual billers. Although organizational NPIs were excluded and alternative handling of ambiguous billers was examined, residual misclassification is possible. Fourth, the market-share trend is not point identified. Fixed membership assigns eventual subspecialty status to pre-certification years, whereas certification-gated classification may assign genuine subspecialty practice to an earlier general category. We therefore present these as plausible scenarios rather than formal bounds.

Approximately 8% of URPS physicians lacked a matched certification year, and urology-pathway URPS status remained fixed because a urology subspecialty certification date was unavailable. Finally, reported line-service units (Tot_Srvcs) may not correspond exactly to unique operations because of claim-line, unit-count, and modifier conventions, and should be interpreted as reported services rather than unique operations.

### Strengths

Strengths include a national sample of publicly reported, nonsuppressed fee-for-service provider-service records; identification of URPS physicians through both certification pathways; explicit separation of organizational and ambiguous billers; within-year and pooled concentration measures with bootstrap confidence intervals; models that account for repeated physician observations; multiple classification sensitivities; and a fully reproducible analytic pipeline.

## Conclusions

Among physician-years observable in fee-for-service Medicare, URPS physicians performed most reported sling services for stress urinary incontinence, and their share increased from 2013 through 2023. The increase was driven primarily by the OB/GYN pathway and declining observable participation among non-URPS clinicians, not by increasing surgeon-level concentration. Because the PUF excludes low-volume physician-years and Medicare Advantage claims and does not distinguish synthetic midurethral from fascial sling procedures, these results describe observable Medicare service patterns rather than the complete national sling market or clinical workforce. All-payer studies linking volume, outcomes, and geographic access are needed before drawing credentialing conclusions.

## References

1. Wu JM. Stress incontinence in women. N Engl J Med. 2021;384(25):2428-2436.
2. Ford AA, Rogerson L, Cody JD, Aluko P, Ogah JA. Mid-urethral sling operations for stress urinary incontinence in women. Cochrane Database Syst Rev. 2017;7(7):CD006375.
3. Kobashi KC, Vasavada S, Bloschichak A, et al. Updates to surgical treatment of female stress urinary incontinence (SUI): AUA/SUFU guideline (2023). J Urol. 2023;209(6):1091-1098.
4. Birkmeyer JD, Stukel TA, Siewers AE, Goodney PP, Wennberg DE, Lucas FL. Surgeon volume and operative mortality in the United States. N Engl J Med. 2003;349(22):2117-2127.
5. Reames BN, Ghaferi AA, Birkmeyer JD, Dimick JB. Hospital volume and operative mortality in the modern era. Ann Surg. 2014;260(2):244-251.
6. American Board of Medical Specialties. ABMS announces approval of new subspecialty certificate in female pelvic medicine and reconstructive surgery. 2011. Available at: https://www.abms.org.
7. Berger AA, Tan-Kim J, Menefee SA. The impact of the 2011 US Food and Drug Administration transvaginal mesh communication on utilization of synthetic mid-urethral sling procedures. Int Urogynecol J. 2021;32(8):2227-2231.
8. Stitzenberg KB, Sigurdson ER, Egleston BL, Starkey RB, Meropol NJ. Centralization of cancer surgery: implications for patient access to optimal care. J Clin Oncol. 2009;27(28):4671-4678.
9. James MB, Theofanides MC, Sui W, Onyeji I, Badalato GM, Chung DE. Sling procedures for the treatment of stress urinary incontinence: comparison of national practice patterns between urologists and gynecologists. J Urol. 2017;198(6):1386-1391.
10. Cantrell AB, Rothschild J, Durbin-Johnson B, Gonzalez R, Kurzrock EA. Surgical trends in the correction of female stress urinary incontinence in academic centers within the United States. Neurourol Urodyn. 2017;36(2):394-398.
11. Rogo-Gupta L, Litwin MS, Saigal CS, Anger JT; Urologic Diseases in America Project. Trends in the surgical management of stress urinary incontinence among female Medicare beneficiaries, 2002-2007. Urology. 2013;82(1):38-41.
12. Siegal AR, Huang Z, Gross MD, Mehraban-Far S, Weissbart SJ, Kim JM. Trends of mesh utilization for stress urinary incontinence before and after the 2011 Food and Drug Administration notification between FPMRS-certified and non-FPMRS-certified physicians: a statewide all-payer database analysis. Urology. 2021;150:151-157.
13. Lee UJ, Feinstein L, Ward JB, et al; Urologic Diseases in America Project. National trends in the surgical management of urinary incontinence among insured women, 2004 to 2013. J Urol. 2020;203(2):365-371.
14. Tabakin AL, Sawhney R, Daily AM, Winkler HA, Shalom DF, Tam J, Lee W. Case log trends of urogynecology and reconstructive pelvic surgery fellows: a comparison of urology- and gynecology-based fellowship programs. Neurourol Urodyn. 2024;43(8):1970-1976.
15. Wang CN, Su IW, Smith AL, Badalato GM, Chung DE. Current exposure to female pelvic medicine and reconstructive surgery faculty during urology residency. Neurourol Urodyn. 2023;42(7):1569-1573.
16. Berger AA, Tan-Kim J, Menefee SA. Surgeon volume and reoperation risk after midurethral sling surgery. Am J Obstet Gynecol. 2019;221(5):523.e1-523.e8.
17. Cartier S, Cerantola GM, Leung AA, Brennand E. The impact of surgeon operative volume on risk of reoperation within 5 years of mid-urethral sling: a systematic review. Int Urogynecol J. 2023;34(5):981-992.
18. Brennand EA, Quan H. Evaluation of the effect of surgeon's operative volume and specialty on likelihood of revision after mesh midurethral sling placement. Obstet Gynecol. 2019;133(6):1099-1108.
19. Holdø B, Svenningsen R. The impact from surgical experience on short- and long-term success rates after mid-urethral sling surgery. Int Urogynecol J. 2026 (epub ahead of print).
20. Gallo K, Weiner H, Mishra K. An update on surgical management for stress urinary incontinence. Curr Opin Obstet Gynecol. 2024;36(6):433-438.

---

## Tables

**Table 1.** Specialty distribution and market-share trends among clinicians billing CPT 57288, Medicare PUF 2013–2023, with URPS split by certification pathway. Δ share is the annual percentage-point change from ordinary least-squares regression on calendar year. Specialty-specific physician counts exceed `r format(v$analytic_physicians, big.mark=",")` because `r v$switchers` physicians changed groups across years. *MIGS estimates are descriptive because only `r v$mig_phys` physicians were identified.

```{r t1}
kt(tab$t1)
```

**Table 2.** Surgeon-level procedural concentration by specialty, based on aggregate provider volume. Raw HHI is not comparable across groups of different size; the normalized HHI (0–1) and effective number of providers are size-adjusted companions.

```{r t2}
kt(tab$t2)
```

**Table 3.** Adjusted rate ratios for annual reported services from a Poisson GEE clustered by NPI (OB/GYN-pathway URPS reference; calendar year centered at 2018; exchangeable correlation; robust standard errors). The first block gives specialty contrasts at mid-study and the 2020 effect; the second block gives each specialty's annual trend as a marginal contrast (the year term plus its specialty-by-year interaction).

```{r t3}
kt(tab$t3)
```

**Table 4.** All-pathway URPS and OB/GYN-based market-share trends under alternative classification scenarios. The scenarios differ in the estimated rate of increase: fixed membership gives the shallower slope because it counts physicians as URPS before certification, and certification-gated classification gives the steeper slope because it removes not-yet-certified physicians from the early URPS count. The 2023 levels converge. These are plausible scenarios rather than formal statistical bounds. All estimated URPS trends are positive and statistically significant.

```{r t4}
kt(tab$t4)
```

---

## Figures

```{r fig1, results='asis'}
show_fig("figure_rate_per_100k.png")
```

**Figure 1.** Utilization of sling surgery for stress urinary incontinence (CPT 57288) in fee-for-service Medicare, 2013–2023. Bars show reported services; the line shows the denominator-adjusted rate per 100,000 female Part B fee-for-service beneficiaries. The rate declined less than the raw count (`r format(v$rate_pct_change, nsmall=1)`% versus `r format(v$count_pct_change, nsmall=1)`%) because the fee-for-service population contracted by `r format(v$den_pct_change, nsmall=1)`%; the linear trend in the rate was not significant (p = `r v$rate_slope_p`).

```{r fig2, results='asis'}
show_fig("figure_market_share.png")
```

**Figure 2.** Market share of sling operations for stress urinary incontinence (CPT 57288) by specialty, fee-for-service Medicare PUF 2013–2023. Stacked areas are the fixed-membership specialty shares; the two black lines are the URPS share under fixed-membership (solid) and certification-gated (dashed) classification. Fixed membership (all-pathway URPS share `r v$fixed_urps_s13`% to `r v$fixed_urps_s23`%; +`r sprintf("%.2f", v$fixed_urps_slope)` percentage points/year) and certification-gated classification (`r v$ct_urps_s13`% to `r v$ct_urps_s23`%; +`r sprintf("%.2f", v$ct_urps_slope)` percentage points/year) differ in the estimated rate of increase and converge by 2023.

```{r fig3, results='asis'}
show_fig("figure_classification_flow.png")
```

**Figure 3.** Provider-classification flow for CPT 57288. Billers are assigned to specialty groups by CMS provider type cross-referenced with the ABOG and ABU rosters. Billers with neither an OB/GYN nor urology CMS type and no ABOG match (`r v$class_reclass_urology` NPIs) are predominantly organizations and non-physician clinicians; organizational NPIs (NPPES entity type 2: ambulatory surgical centers, hospitals, laboratories; `r v$class_excluded_facility`) are excluded and the remaining individual clinicians (`r v$other_phys`) form an Other/uncertain group, rather than being counted as urologists. Specialty-specific counts exceed the analytic cohort because some physicians changed groups across years.

---

## Supplementary Information

Supplementary tables are generated by pipeline step 06 from the same frozen cache as the main results, so they remain consistent with the numbers above.

**Supplementary Table S1.** Annual procedural concentration by calendar year and specialty. Total procedures, observable surgeons, median annual volume, Gini coefficient, HHI, and top-decile, top-quintile, and bottom-half shares are reported for every year. Within-year concentration was low and stable across the study period.

```{r s1}
s <- read_supp("table_5_annual_concentration")
if (!is.null(s)) kt(s) else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S2.** Trend regressions for annual concentration and volume measures on calendar year, overall and by specialty (ordinary least squares).

```{r s2}
s <- read_supp("table_6_concentration_trends")
if (!is.null(s)) {
  s <- dplyr::mutate(s,
    `Slope / year` = round(as.numeric(`Slope / year`), 3),
    `R-squared`    = round(as.numeric(`R-squared`), 3),
    `p-value`      = fmt_pcol(`p-value`))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S3.** Specialty-specific market-share trends. Each specialty's annual share of observed procedures is regressed on calendar year separately (fixed-membership classification).

```{r s3}
s <- read_supp("table_12_specialty_share_trends")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Specialty       = specialty_group,
    `2013 share (%)`= share_start,
    `2023 share (%)`= share_end,
    `Slope (pp/yr)` = round(as.numeric(slope_pp_yr), 3),
    `p-value`       = fmt_pcol(p_value))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S4.** Adjusted rate ratios for annual sling volume from a Poisson GEE clustered by NPI, excluding calendar year 2020 (COVID-19 sensitivity analysis; URPS reference).

```{r s4}
s <- read_supp("table_8b_volume_gee_no2020")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Term          = gsub("specialty_group", "Specialty: ", term),
    `RR (95% CI)` = `RR (95% CI)`,
    `p-value`     = fmt_pcol(p_value))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S5.** Per-physician volume comparisons (one median value per physician), Kruskal-Wallis and pairwise Wilcoxon tests with Bonferroni correction.

```{r s5}
s <- read_supp("table_9_per_physician")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Test        = test,
    Statistic   = ifelse(is.na(suppressWarnings(as.numeric(statistic))), "n/a",
                         sprintf("%.2f", suppressWarnings(as.numeric(statistic)))),
    df          = df,
    `p-value`   = p_formatted)
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S6a.** Specialty distribution under alternative classification schemes (time-varying, modal, ever-URPS/MIGS, certification-gated).

```{r s6a}
s <- read_supp("table_10_classification_sensitivity")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Scheme       = scheme,
    Specialty    = specialty_group,
    `N providers`= n_providers,
    Procedures   = procedures,
    `% of all`   = sprintf("%.1f%%", as.numeric(pct_of_all)))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S6b.** All-pathway URPS and OB/GYN-based market-share trends under each classification scheme. The URPS increase remains positive and significant in every scheme.

```{r s6b}
s <- read_supp("table_10b_classification_trends")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    Scheme               = scheme,
    `URPS slope (pp/yr)` = round(as.numeric(urps_slope_pp_yr), 3),
    `URPS p`             = fmt_pcol(urps_p),
    `Gyn slope (pp/yr)`  = round(as.numeric(gyn_slope_pp_yr), 3),
    `Gyn p`              = fmt_pcol(gyn_p))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S7.** Observable surgeons and URPS share by provider practice state (geography, secondary analysis). A state with no observable URPS surgeon may still have URPS surgeons who each fell below the CMS reporting threshold.

```{r s7}
s <- read_supp("table_13_geography_by_state")
if (!is.null(s)) {
  s <- dplyr::transmute(s,
    State                 = state,
    `Observable surgeons` = n_observable_surgeons,
    `URPS surgeons`       = n_urps,
    `URPS share`          = sprintf("%.1f%%", as.numeric(urps_share_pct)),
    `Observable URPS?`    = ifelse(as.logical(has_observable_urps), "Yes", "No"))
  kt(s)
} else cat("_Run pipeline step 06 to generate this table._")
```

**Supplementary Table S8.** Suppression illustration (not an empirically estimated sensitivity). Hypothetical suppressed low-volume clinicians (25% and 50% of the observed count, each performing 1–10 services) were added to each group before recomputing the Gini coefficient and HHI. Adding these clinicians increased estimated inequality, particularly in the 50% scenario, while HHI remained low. Because neither the number nor the volume distribution of suppressed clinicians is known, these illustrations do not identify full-market concentration; they support only the narrower conclusion that observable services were not dominated by a handful of surgeons.

```{r s8}
if (!is.null(tab$t_suppress)) kt(tab$t_suppress) else cat("_Not available._")
```

**Supplementary Table S9.** Annual observable participation transitions using a two-year washout. Newly observable surgeons were absent in both prior observable years; surgeons no longer observable were absent in both subsequent observable years. Newly observable and continuing counts are undefined for the first two years, and no-longer-observable counts are undefined for the last two years. These describe threshold crossings, not definitive entry into or exit from practice.

```{r s9}
if (!is.null(tab$t5)) kt(tab$t5) else cat("_Not available._")
```

Newly observable episodes by specialty: URPS, `r v$ent_urps`; other OB/GYN, `r v$ent_gob`; non-URPS urology, `r v$ent_uro`; MIGS, `r v$ent_migs`.

**Supplementary Table S11.** Classification sensitivity for ambiguous and facility billers (reviewer concern). The primary analysis excludes facility NPIs (NPPES entity type 2) and the remaining non-physician or unclassifiable billers from the physician cohort. Retaining those billers as a separate "Other/uncertain" group changes the well-populated specialty shares only slightly, whereas assigning them to urology (the legacy approach) roughly doubles the apparent non-URPS urology share. The URPS increase is significant under all three.

```{r s11}
if (!is.null(tab$t_classif_sens)) kt(tab$t_classif_sens) else cat("_Not available._")
```

---

## Supplementary Figures

```{r sfig1, results='asis'}
show_fig("figure_2_volume_distribution.png")
```

**Supplementary Figure S1.** Annual sling service volume (CPT 57288) by specialty (violin and box plots; logarithmic scale). The minimum observable volume is 11 because of CMS cell suppression.

```{r sfig2, results='asis'}
show_fig("figure_3_lorenz_curve.png")
```

**Supplementary Figure S2.** Lorenz curves of procedural concentration by specialty. Greater distance from the diagonal indicates greater concentration.

```{r sfig3, results='asis'}
show_fig("figure_4_concentration_trends.png")
```

**Supplementary Figure S3.** Annual within-year concentration by specialty (Gini, HHI, top-20% share, bottom-50% share). Within-year concentration was low and stable. MIGS was excluded because too few surgeons were observed each year for stable estimates.

```{r sfig4, results='asis'}
show_fig("figure_5_supply_trends.png")
```

**Supplementary Figure S4.** Observable surgeons and total reported services by year and specialty. The decline in the observable surgeon pool was concentrated in urology and General OB/GYN.
